---
title: "Political Geography"
author: "Alex Nelson"
date: "7/25/2019"
output:
  md_document:
    variant: gfm
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
library(ape)
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(maps)
library(usmap)

library(MXM) # for transitiveClosure()
library(zoo)

library(rmarkdown)

library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```
```{r, include=FALSE}
source("../R/states.R")
```

```{r}
foo <- load_obj(precinct_2016_path)
foo %>% group_by(mode) %>% summarize(count = n())
```

```{r}
pres <- load_obj(county_path)
```


```{r}
bar <- foo %>%
  rename(FIPS = county_fips) %>%
  group_by(FIPS, mode, county_name, state) %>%
  summarize(votes = sum(votes)) %>%
  inner_join(pres %>%
              filter(year == 2016) %>%
              group_by(FIPS) %>%
              summarize(totalvotes0 = sum(candidatevotes),
                        tv = max(totalvotes)),
            by="FIPS") %>%
  mutate(diff = tv - votes)
```

```{r}
missing_votes <- bar %>%
  filter(votes == tv) %>%
  group_by(state) %>%
  summarize(counts = n()) %>%
  arrange(-counts) %>%
  inner_join(foo %>%
               select(state, county_name) %>%
               unique %>%
               group_by(state) %>%
               summarize(num_counties = n()), by="state") %>%
  mutate(diff = num_counties - counts) %>%
  arrange(-diff)

```


```{r}

kable(missing_votes)

```



# Introduction

The goal is to try to cluster states based on geographic proximity and similar voting tendency. The first thing we do is load the data.

```{r}
load("../data/elections/senate/1976_2018_senate.RData")
senate_data <- x %>%
  filter(stage == "gen") %>%
  mutate(party = ifelse((party == "democrat" | party == "democratic-farmer-labor") & writein == F, "democrat",
                        ifelse(party == "republican" & writein == F, party, "$third")))

load("../data/elections/house/1976_2018_house.RData")
house_data <- x %>%
  filter(stage == "gen") %>%
  mutate(party = ifelse((party == "democrat" | party == "democratic-farmer-labor") & writein == F, "democrat",
                        ifelse(party == "republican" & writein == F, party, "$third")))

load("../data/elections/presidential/state/1976_2016_president.RData")
president_data <- x %>%
  mutate(party = ifelse((party == "democrat" | party == "democratic-farmer-labor") & writein == F, "democrat",
                        ifelse(party == "republican" & writein == F, party, "$third"))) %>%
  group_by(state, year) %>%
  mutate(totalvotes = sum(candidatevotes)) %>%
  ungroup
```

```{r}
max_party <- function(df0) {
  (df0 %>% arrange(-candidatevotes) %>% head(n=1))[["party"]]
}

foo_2002 <- house_data %>%
  filter(year == 2002,
         # runoff == FALSE,
         special == FALSE) %>%
  group_by(state, district) %>%
  nest() %>%
  mutate(winner = lapply(data, max_party)) %>%
  unnest(winner) %>%
  ungroup() %>%
  select(-data) %>%
  group_by(state, winner) %>%
  summarize(count = n()) %>%
  spread(winner, count, fill = 0) %>%
  group_by(state) %>%
  mutate(electoral = ifelse(republican > democrat,
                            2 + republican + democrat,
                            ifelse(republican < democrat,
                                   -2 - republican - democrat,
                                   0))) %>%
  ungroup()
```


We now need to work our way to produce a triple for each state.

```{r}
house <- house_data %>%
  filter(special == F) %>%
  group_by(state, party, district, year) %>%
  summarize(candidatevotes = sum(candidatevotes),
            totalvotes = totalvotes[[1]]) %>%
  spread(party, candidatevotes, fill = 0) %>%
  select(-`<NA>`) %>%
  mutate(third = `$third`/totalvotes,
         dem = democrat/totalvotes,
         rep = republican/totalvotes)

senate <- senate_data %>%
  filter(special == F) %>%
  group_by(state, year, party, totalvotes) %>%
  summarize(candidatevotes = sum(candidatevotes)) %>%
  ungroup() %>%
  spread(party, candidatevotes, fill = 0) %>%
  select(-`<NA>`) %>%
  mutate(third = `$third`/totalvotes,
         dem = democrat/totalvotes,
         rep = republican/totalvotes)

president <- president_data %>% 
  group_by(state, year, party, totalvotes) %>%
  summarize(candidatevotes = sum(candidatevotes)) %>%
  ungroup() %>%
  spread(party, candidatevotes, fill = 0) %>%
  select(-`<NA>`) %>%
  mutate(third = `$third`/totalvotes,
         dem = democrat/totalvotes,
         rep = republican/totalvotes)
```

We create a vector for a given state representing the `third`, `dem`, `rep` percents ordered in time. Then these vectors are used to compute the correlations between states, pair-wise.

```{r}
state_to_vec <- function(data, state_name) {
  (data %>%
    filter(state == state_name) %>%
    gather("party", "percent", third, dem, rep) %>%
    select(state, year, party, percent) %>%
    arrange(year, party))$percent
}
```

```{r}
states <- factor(sort(unique(president$state)))
state_vecs <- list()
for (state in states) {
  # state_vecs <- append(state_vecs, list(state = state_to_vec(president, state)))
  state_vecs[[as.character(state)]] <- state_to_vec(president, state)
}
```

```{r}
cor_mat <- data.frame(from = c(),
                      to = c(),
                      val = c())
cor_mat_spearman <- data.frame(from = c(),
                      to = c(),
                      val = c())
for (i in 1:50) {
  from <- states[i]
  for (j in (1 + i):51) {
    to <- states[j]
    
    val <- cor(state_vecs[[i]], state_vecs[[j]], method = "spearman")
    cor_mat_spearman <- rbind(cor_mat_spearman, data.frame(from = from,
                                         to = to,
                                         val = val))
    
    val <- cor(state_vecs[[i]], state_vecs[[j]])
    cor_mat <- rbind(cor_mat, data.frame(from = from,
                                         to = to,
                                         val = val))
  }
}
```

```{r, spearman_histogram}
ggplot(cor_mat_spearman, aes(x=val)) + 
  geom_histogram(color="black", fill="white")
```

```{r, pearson_histogram}
ggplot(cor_mat, aes(x=val)) + 
  geom_histogram(color="black", fill="white")
```

## Grouping Pair-wise Correlations

```{r}

```


```{r}
mat <- matrix(0, 51, 51)
edges <- filter(cor_mat, val > quantile(cor_mat$val, probs = c(0.925))[1])
for (i in 1:51) {
  relevant_edges <- rbind(filter(edges, from == states[i], to %in% neighboring_states[[states[i]]]),
                          filter(edges, to == states[i], from %in% neighboring_states[[states[i]]]))
  
  for (j in relevant_edges$from) {
    to <- factor(j, states)
    mat[i, as.integer(to)] <- 1
    mat[as.integer(to), i] <- 1
  }
  for (j in relevant_edges$to) {
    to <- factor(j, states)
    mat[i, as.integer(to)] <- 1
    mat[as.integer(to), i] <- 1
  }
}
rownames(mat) <- states
colnames(mat) <- states

tr_ad <- transitiveClosure(mat)
rownames(tr_ad) <- states
colnames(tr_ad) <- states
tr_ad0 <- tr_ad

for (i in 1:51) {
  if (all(tr_ad[i,] == 0) & length(neighboring_states[[states[i]]]) > 0) {
    print(i)
    choices <- rbind(filter(cor_mat, from == states[i], to %in% neighboring_states[[states[i]]]),
                     filter(cor_mat, to == states[i], from %in% neighboring_states[[states[i]]]))
    to <- (choices %>% arrange(-val))$to[1]
    print(head((choices %>% arrange(-val))))
    if (to == states[i]) {
      to <- factor((choices %>% arrange(-val))$from[1], states)
    } else {
      to <- factor(to, states)
    }
    tr_ad[i, as.integer(to)] <- 1
    tr_ad[as.integer(to), i] <- 1    
  }
}

tr_ad <- transitiveClosure(tr_ad)
rownames(tr_ad) <- states
colnames(tr_ad) <- states

ci <- 1
clusters <- data.frame(region = c(),
                       cluster = c())
for (i in 1:51) {
  others <- rownames(as.matrix(Filter(function(x) {x > 0}, tr_ad[i, ])))
  if (length(others) > 0 & !(tolower(as.character(states[i])) %in% clusters$region)) {
    for(other in sort(append(others, as.character(states[i])))) {
      clusters <- rbind(clusters, data.frame(region = c(tolower(as.character(other))),
                                             cluster = c(ci)))
    }
    ci <- ci + 1
  }
}
clusters$cluster <- factor(clusters$cluster)

us_states <- map_data("state")
clustered_states <- us_states %>%
  inner_join(clusters, by = "region")
```


```{r, clustered_map}
red <- "#FF2700"
orange <- "#ff7400"
green <- "#77AB43"
blue <- "#008FD5"
party_colors <- c(blue, # 1
                  orange, # 2
                  green, # 3
                  blue, # 4
                  green, # 5
                  red, # 6
                  blue, # 7 
                  orange, # 8
                  green, # 9
                  red, # 10
                  red, # 11
                  blue) # 12

ggplot(data = clustered_states,
            mapping = aes(x = long, y = lat, 
                          fill = cluster, group = group)) +
  geom_polygon(color = "gray90", size = 0.1) + 
  coord_fixed(1.3) +
  scale_fill_manual(values = party_colors) +
  guides(fill = FALSE) + # no legend!
  labs(caption="http://political-arithmetic.blogspot.com",
       title= "State Clusters", 
       subtitle="Clustering states based on Presidential voting behavior 1976-2016")
```


# Using Spatial Autocorrelation, Within a State

Within a given state, we want to cluster adjacent counties together based on presidential voting patterns.

```{r}
source("../R/presidential_elections.R")

county_data <- make_party_into_factor(load_obj(county_path)) %>%
  group_by(year, FIPS, party, candidate) %>%
  mutate(percent = sum(candidatevotes)/sum(totalvotes)) %>%
  ungroup()
```

## Neighboring Counties Data

We can download a CSV from [NBER](https://www.nber.org/data/county-adjacency.html)
to track neighboring counties.

```{r}
neighboring_counties <- read_csv("../data/census/county_adjacency2010.csv",
                                 col_types = cols(countyname   = col_character(),
                                                  fipscounty   = col_integer(),
                                                  neighborname = col_character(),
                                                  fipsneighbor = col_integer())) %>%
  select(fipscounty, fipsneighbor) %>%
  filter(fipscounty != fipsneighbor) %>%
  rename(from = fipscounty,
         to = fipsneighbor)
```

## Test Case: Iowa

Lets try seeing what happens for Iowa based on 2012.

```{r}
ia_votes <- (county_data %>%
               filter(state == "Iowa",
                      year == 2016,
                      party == "republican") %>%
               select(FIPS, percent) %>%
               arrange(FIPS))$percent
```

We need to compute the adjacency matrix.

```{r}
ia_fips <- sort(unique((county_data %>%
               filter(state == "Iowa",
                      year == 2016,
                      party == "republican") %>%
               select(FIPS, percent))$FIPS))

ia_neighboring_counties <- neighboring_counties %>%
  filter(from %in% ia_fips, to %in% ia_fips)

ia_adj_mat <- matrix(0, nrow = length(ia_fips), ncol = length(ia_fips))
ia_laplace <- matrix(0, nrow = length(ia_fips), ncol = length(ia_fips))

for (i in 1:length(ia_fips)) {
  county_i <- ia_fips[[i]]
  neighbors <- filter(ia_neighboring_counties, from == county_i)
  for (j in 1:length(ia_fips)) {
    county_j <- ia_fips[[j]]
    if (county_j %in% neighbors$to && !(i == j)) {
      ia_adj_mat[i,j] <- 1
    }
  }
  ia_laplace[i,i] <- sum(ia_adj_mat[i, ])
}
ia_laplace <- ia_laplace - ia_adj_mat
```

Then we compute Moran's I matrix:

```{r}
ia_moran <- rep(0, length(ia_votes))

m2 <- sd(ia_votes)
mu <- mean(ia_votes)
for (i in 1:length(ia_moran)) {
  ia_moran[[i]] <- ((ia_votes[[i]] - mu)/m2)*sum(ia_votes[[i]] * (ia_adj_mat %*% ia_votes)[[i]])
}

ia_results <- data.frame(fips = ia_fips,
                         i = ia_moran)
```

We plot it out:

```{r, ia-test1}
plot_usmap(data = ia_results, regions = "counties", include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "blue", high = "red", name = "Vote Homogeneity", label = scales::comma
  ) + theme(legend.position = "right")
```

## Laplacian Graph

For Iowa, I'm just curious what the Laplacian's eigenvectors look like when graphed.

```{r}
ia_eigen <- eigen(ia_laplace)
v2 <- ia_eigen$vectors
v2[which(v2 < round(mean(v2), digits = 2))] <- 0
```

The Fourier transformed version of the `ia_votes` vector is just the matrix multiplication of the eigenvectors with the `ia_votes`:

```{r, ia-fourier-votes}
ia_results <- data.frame(fips = ia_fips,
                         i = ia_eigen$vectors %*% ia_votes)
plot_usmap(data = ia_results, regions = "counties", include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "blue", high = "red", name = "Fourier Transformed Votes", label = scales::number
  ) + theme(legend.position = "right")
```


```{r, ia-eigenvec1}
ia_results <- data.frame(fips = ia_fips,
                         i = v2[,1])
plot_usmap(data = ia_results, regions = "counties", include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Eigenvector", label = scales::comma
  ) + theme(legend.position = "right")
```

```{r, ia-eigenvec2}
ia_results <- data.frame(fips = ia_fips,
                         i = v2[,2])
plot_usmap(data = ia_results, regions = "counties", include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Eigenvector", label = scales::comma
  ) + theme(legend.position = "right")
```
```{r, ia-eigenvec3}
ia_results <- data.frame(fips = ia_fips,
                         i = v2[,3])
plot_usmap(data = ia_results, regions = "counties", include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Eigenvector", label = scales::comma
  ) + theme(legend.position = "right")
```
```{r, ia-eigenvec4}
ia_results <- data.frame(fips = ia_fips,
                         i = v2[,4])
plot_usmap(data = ia_results, regions = "counties", include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Eigenvector", label = scales::comma
  ) + theme(legend.position = "right")
```

```{r, ia-eigenvec5}
ia_results <- data.frame(fips = ia_fips,
                         i = v2[,5])
plot_usmap(data = ia_results, regions = "counties", include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Eigenvector", label = scales::comma
  ) + theme(legend.position = "right")
```

```{r, ia-eigenvec6}
ia_results <- data.frame(fips = ia_fips,
                         i = v2[,6])
plot_usmap(data = ia_results, regions = "counties", include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Eigenvector", label = scales::comma
  ) + theme(legend.position = "right")
```
```{r, ia-eigenvec7}
ia_results <- data.frame(fips = ia_fips,
                         i = v2[,7])
plot_usmap(data = ia_results, regions = "counties", include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Eigenvector", label = scales::comma
  ) + theme(legend.position = "right")
```
```{r, ia-eigenvec8}
ia_results <- data.frame(fips = ia_fips,
                         i = v2[,8])
plot_usmap(data = ia_results, regions = "counties", include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Eigenvector", label = scales::comma
  ) + theme(legend.position = "right")
```
```{r, ia-eigenvec9}
ia_results <- data.frame(fips = ia_fips,
                         i = v2[,9])
plot_usmap(data = ia_results, regions = "counties", include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Eigenvector", label = scales::comma
  ) + theme(legend.position = "right")
```
```{r, ia-eigenvec10}
ia_results <- data.frame(fips = ia_fips,
                         i = v2[,10])
plot_usmap(data = ia_results, regions = "counties", include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Eigenvector", label = scales::comma
  ) + theme(legend.position = "right")
```
```{r, ia-eigenvec11}
ia_results <- data.frame(fips = ia_fips,
                         i = v2[,11])
plot_usmap(data = ia_results, regions = "counties", include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Eigenvector", label = scales::comma
  ) + theme(legend.position = "right")
```
## Wrap this in a function

We first have a helper function to construct the adjacency matrix for counties within a given state:

```{r}
make_county_fips <- function(state_fullname) {
  sort(unique((county_data %>%
                 filter(state == state_fullname,
                        party == "republican") %>%
                 select(FIPS, percent))$FIPS))
}

make_county_adj_mat <- function(state_fullname) {
  county_fips <- make_county_fips(state_fullname)
  
  state_neighboring_counties <- neighboring_counties %>%
    filter(from %in% county_fips, to %in% county_fips)
  
  county_adj_mat <- matrix(0, nrow = length(county_fips), ncol = length(county_fips))
  for (i in 1:length(county_fips)) {
    county_i <- county_fips[[i]]
    neighbors <- filter(state_neighboring_counties, from == county_i)
    for (j in 1:length(county_fips)) {
      county_j <- county_fips[[j]]
      
      if ((county_j %in% neighbors$to) && !(i == j)) {
        county_adj_mat[i, j] <- 1
      }
    }
  }
  diag(county_adj_mat) <- 0
  return(county_adj_mat)
}
```


We can now wrap this in a function, asking for a year and state to plot out. We have two versions: one for using a weighted average (weighing recent elections heavier), and one for a "vanilla" average.

Moran's I will tell us the "local variance" relative to the "global variance" ratio, for each county.

```{r}
plot_moran_i <- function(state_fullname, use_weighted_avg = F) {
  if (use_weighted_avg) {
    print("In the 'use weighted avg' branch...")
    state_votes <- (county_data  %>%
                       filter(state == state_fullname,
                              party == "republican") %>%
                       # group_by(year) %>%
                       mutate(w_cv = ((year - 1996)/4)*candidatevotes,
                              w_tv = ((year - 1996)/4)*totalvotes) %>%
                       # ungroup() %>%
                       group_by(FIPS) %>%
                       summarize(percent = sum(w_cv)/sum(w_tv)) %>%
                       arrange(FIPS))$percent
    mu <- (county_data %>%
             filter(state == state_fullname,
                    party == "republican") %>%
             group_by(year) %>%
             mutate(w_cv = ((year - 1996)/4)*candidatevotes,
                    w_tv = ((year - 1996)/4)*totalvotes) %>%
             ungroup() %>%
             summarize(mu = sum(w_cv)/sum(w_tv)))$mu
  } else {
    state_votes <- (county_data  %>%
                       filter(state == state_fullname,
                              party == "republican") %>%
                       group_by(FIPS) %>%
                       summarize(percent = sum(candidatevotes)/sum(totalvotes)) %>%
                       arrange(FIPS))$percent
    mu <- (county_data %>%
             filter(state == state_fullname,
                    party == "republican") %>%
             summarize(mu = sum(candidatevotes)/sum(totalvotes)))$mu
  }
  print(paste0("mu = ", mu))
  
  county_adj_mat <- make_county_adj_mat(state_fullname)
  
  print("Computing Moran's I vector")
  state_moran <- rep(0, length(state_votes))

  S0 <- sum(unlist(county_adj_mat))
  z <- state_votes - mu
  m2 <- sum(z**2)/(length(z))
  v <- (county_adj_mat %*% z)
  
  state_moran <- (z*v)/(m2)
  
  county_fips <- make_county_fips(state_fullname)
  results <<- data.frame(fips = county_fips,
                         i = state_moran)
  
  plot_usmap(data = results,
             regions = "counties",
             include = state_fullname,
             values = "i") + 
    scale_fill_continuous(
      low = "white", high = "red", name = "Vote Homogeneity") + theme(legend.position = "right")
}
```
462
```{r, ia-moran-unweighted}
plot_moran_i("Iowa", F)
```

```{r, ia-moran-weighted}
plot_moran_i("Iowa", T)
```
## Local Geary

The local Geary's C tells us how similar a county is to its neighbors, intuitively the off-diagonal rows of the cross-covariance matrix.

```{r}
plot_geary <- function(state_fullname) {
  state_votes <- (county_data %>%
                    filter(state == state_fullname,
                           party == "republican") %>%
                    group_by(FIPS) %>%
                    summarize(percent = sum(candidatevotes)/sum(totalvotes)) %>%
                    arrange(FIPS))$percent
  
  county_adj_mat <- make_county_adj_mat(state_fullname)
  
  mu <- (county_data %>%
           filter(state == state_fullname,
                  party == "republican") %>%
           summarize(mu = sum(candidatevotes)/sum(totalvotes)))$mu

  S0 <- sum(unlist(county_adj_mat))
  # m2 <- sum((state_votes - mu)**2)/(length(state_votes))
  z <- state_votes - mu
  m2 <- sum(z**2)/(length(z) - 1)

  geary <- rep(0, length(z))
  for (i in 1:length(geary)) {
    u <- (z[[i]] - z)**2
    geary[[i]] <- (county_adj_mat[i,] %*% u)[[1]]
  }
  geary <<- geary/(2*sum(unlist(county_adj_mat))*m2)
  
  county_fips <- sort(unique((county_data %>%
                                filter(state == state_fullname,
                                       party == "republican") %>%
                                select(FIPS, percent))$FIPS))
  results <<- data.frame(fips = county_fips,
                         i = geary)
  
  plot_usmap(data = results,
             regions = "counties",
             include = state_fullname,
             values = "i") + 
    scale_fill_continuous(
      low = "white", high = "red", name = "Vote Homogeneity") + theme(legend.position = "right")
}
```

For example, Ohio highlights major cluster centers.

```{r, ia-geary}
plot_geary("Ohio")
```


## Local Lags

We can plot the percent a county voted for the Republican versus the average percent of its neighbors.

```{r}
plot_lag <- function(state_fullname) {
  state_votes <- (county_data %>%
                    filter(state == state_fullname,
                           party == "republican") %>%
                    group_by(year) %>%
                    mutate(w_cv = (((year - 1996)/4))*candidatevotes,
                           w_tv = (((year - 1996)/4))*totalvotes) %>%
                    ungroup() %>%
                    group_by(FIPS) %>%
                    summarize(percent = sum(w_cv)/sum(w_tv)) %>%
                    arrange(FIPS))$percent
  
  county_fips <- sort(unique((county_data %>%
                                filter(state == state_fullname,
                                       party == "republican") %>%
                                select(FIPS, percent))$FIPS))
  
  county_adj_mat <- make_county_adj_mat(state_fullname)
  for(i in 1:nrow(county_adj_mat)) {
    county_adj_mat[i, ] <- county_adj_mat[i, ]/sum(county_adj_mat[i, ])
  }

  mu <- (county_data %>%
           filter(state == state_fullname,
                  party == "republican") %>%
           mutate(w_cv = (((year - 1996)/4))*candidatevotes,
                  w_tv = (((year - 1996)/4))*totalvotes) %>%
           summarize(mu = sum(w_cv)/sum(w_tv)))$mu
  z <- state_votes
  neighbor_lag <- (county_adj_mat %*% z)
  results <- data.frame(county = z,
                        neighbor = neighbor_lag)
  ggplot(results, aes(x = county, y = neighbor)) + 
    geom_point(shape = 18, color = "blue") +
    geom_smooth(method = lm,  linetype = "dashed",
                color = "darkred", fill = "blue") +
    geom_abline(intercept = 0, slope = 1) +
    labs(title = paste0("County Presidential Voting in ", state_fullname),
         subtitle = "Compared to average neighboring counties") +
    ylab("Average neighbor's vote for Republican (%)") +
    xlab("County Republican Vote (%)")
}
```

Example:

```{r, lag-ia}
plot_lag("Iowa")
```


```{r, lag-pa}
plot_lag("Pennsylvania")
```

```{r, lag-oh}
plot_lag("Ohio")
```

```{r, lag-mi}
plot_lag("Michigan")
```

# Using differences between Democrats and Republicans

Consider now using the difference in vote percentage between the democratic candidate and the republican candidate.

```{r}
delta_data <- county_data %>%
  filter(party %in% c("democrat", "republican")) %>%
  select(-candidate) %>%
  spread(party, candidatevotes, fill = 0) %>%
  group_by(year) %>%
  mutate(nation_dem = sum(democrat),
         nation_rep = sum(republican)) %>%
  ungroup() %>%
  group_by(state,county,year,FIPS) %>%
  summarize(republican = sum(republican),
            democrat = sum(democrat),
            nation_dem = max(nation_dem),
            nation_rep = max(nation_rep),
            percent = (democrat - republican)/(democrat + republican))
```



```{r}
ia_votes <- (delta_data %>%
               filter(year == 2016,
                      state == "Iowa") %>%
               arrange(FIPS))$percent

ia_moran <- rep(0, length(ia_votes))

mu <- (delta_data %>%
         filter(year == 2016) %>%
         mutate(mu = (nation_dem - nation_rep)/(nation_dem + nation_rep)))$mu[[1]]
m2 <- sum((ia_votes - mu)**2)/(length(ia_votes) - 1)


m2 <- sd(ia_votes)
mu <- mean(ia_votes)
for (i in 1:length(ia_moran)) {
  ia_moran[[i]] <- ((ia_votes[[i]] - mu)/m2)*sum(ia_votes[[i]] * (ia_adj_mat %*% ia_votes)[[i]])
}

ia_results <- data.frame(fips = ia_fips,
                         i = ia_moran)
```

```{r, ia-test1}
plot_usmap(data = ia_results,
           regions = "counties",
           include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Vote Homogeneity", label = scales::comma
  ) + theme(legend.position = "right")
```

```{r, ia-test1}
plot_usmap(data = ia_results,
           regions = "counties",
           include = "IA", values = "i") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Vote Homogeneity", label = scales::comma
  ) + theme(legend.position = "right")
```

```{r}
ia_results <- data.frame(percent = ia_percent,
                         neighbor = (ia_adj_mat * (ia_neighboring_counties %>% group_by(from) %>% summarize(count = 1/n()) %>% arrange(from))$count) %*% ia_percent)
ggplot(ia_results, aes(x = percent, y = neighbor)) + 
  geom_point(shape = 18, color = "blue") +
  geom_smooth(method = lm,  linetype = "dashed",
              color = "darkred", fill = "blue") +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "County Presidential Voting in Iowa",
       subtitle = "Compared to average neighboring counties") +
  ylab("Average neighbor's vote (%)") +
  xlab("County Vote (%)")
```