---
title: "Democratic Primaries"
author: "Alex Nelson"
date: "6/19/2019"
output:
  md_document:
    variant: gfm
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
library(rvest)                    # scraping
library(tidyverse)                # data cleaning and ggplot2 - tidy!
library(stringr)                  # string manipulation
library(tibble)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
source("../R/states.R")
```

REMEMBER: the McGovern-Fraser reforms of 1972 didn't take effect until the 1976 cycle. So there is no relevant data BEFORE 1976.

# Cleaning up the Primary Data

I'd like to organize a tidy primary data set, I'm consulting [US Election Atlas](https://uselectionatlas.org/RESULTS/national.php?year=1992&off=0&elect=1&f=0) and wikipedia.
Wikipedia provides clearer data for the 1976-2004 primaries, but sadly for 2008 and 2016 wikipedia falls down short.

## Helper Functions

We have a few helper functions to classify each contest as a primary or a caucus. Some states had both a primary and a caucus (Washington state, apparently, had both until the 1990s).

```{r}
is_caucus <- function(state) {
  grepl("caucus", state, ignore.case=T)
}
is_primary <- function(state) {
  grepl("primary", state, ignore.case=T)
}
remove_tail <- function(s) {
  sub("\\s*\\[.*\\].*$", "", s)
}
cleanup_state <- function(state) {
  sub("\\s*[\\(]?caucus[es\\)]{0,3}.*", "",
      sub("\\s*[\\(]?primary[\\)]?.*", "", state, ignore.case=T),
      ignore.case=T)
}
```

## 1976 Primary

```{r}
tidy_1976 <- function() {
  html("../data/primary/wiki_1976.html") %>%
  html_table(header=T) %>%
  `[[`(1) %>%
  gather(`Jimmy Carter`,`Jerry Brown`,`George Wallace`,
         `Morris Udall`,`Scoop Jackson`,`Frank Church`,
         `Lloyd Bentsen`,`Robert Byrd`,`Sargent Shriver`,
         `Fred Harris`,`Ellen McCormack`,`Milton Shapp`,
         `Birch Bayh`,`Uncommitted`,
         key = "Candidate", value = "vote_perc") %>%
  mutate(Date = as.Date(paste0(Date, " 1976"), format = "%B %d %Y"),
         vote_perc = ifelse(vote_perc == "-", "0%", vote_perc),
         vote_perc = as.numeric(gsub("%.*$", "", vote_perc)),
         vote_perc = ifelse(is.na(vote_perc), 0, vote_perc)) %>%
  rename(date = Date,
         state = Contest,
         candidate = Candidate) %>%
  group_by(state) %>%
  mutate(won_state = (vote_perc == max(vote_perc, na.rm = T)),
         is_nominee = (candidate == "Jimmy Carter")) %>%
  ungroup() %>%
  mutate(event_type = ifelse(is_caucus(state), "caucus", "primary"),
         state = remove_tail(cleanup_state(state)))
}
write_csv(tidy_1976(), "../data/primary/democrat_1976.csv")
```
## 1980 Primary

```{r}
tidy_1980 <- function() {
  html("./data/primary/wiki_1980.html") %>%
    html_table(header = T) %>%
    `[[`(1) %>%
    gather(`Jimmy Carter`,`Ted Kennedy`,`Jerry Brown`,
           `Lyndon LaRouche`,`Cliff Finch`,`Richard Kay`,
           `Unpledged`,`Others`,
           key = "Candidate", value = "vote_perc") %>%
    mutate(Date = as.Date(paste0(Date, " 1980"), format = "%B %d %Y"),
           vote_perc = ifelse(vote_perc == "-", "0%", vote_perc),
           vote_perc = as.numeric(gsub("%.*$", "", vote_perc)),
           vote_perc = ifelse(is.na(vote_perc), 0, vote_perc)) %>%
    rename(date = Date,
           state = State,
           candidate = Candidate) %>%
    group_by(state) %>%
    mutate(won_state = (vote_perc == max(vote_perc, na.rm = T)),
           is_nominee = (candidate == "Jimmy Carter")) %>%
    ungroup() %>%
    mutate(event_type = ifelse(is_caucus(state), "caucus", "primary"),
           state = remove_tail(cleanup_state(state)))
} 
write_csv(tidy_1980(), "../data/primary/democrat_1980.csv")
```



## 1984 Primary


```{r}
tidy_1984 <- function() {
  html("../data/primary/wiki_1984.html") %>%
    html_table(header = T) %>%
    `[[`(1) %>%
    gather(`Walter Mondale`,`Gary Hart`,`Jesse Jackson`,`John Glenn`,
           `George McGovern`,`Ernest Hollings`,`Reubin Askew`,
           `Alan Cranston`,`Uncommitted`,`Lyndon LaRouche`,
           key = "Candidate", value = "vote_perc") %>%
    mutate(Date = as.Date(paste0(Date, " 1984"), format="%B %d %Y"),
           vote_perc = as.numeric(gsub("%", "", vote_perc)),
           vote_perc = ifelse(is.na(vote_perc), 0, vote_perc)) %>%
    rename(date = Date,
           state = State,
           candidate = Candidate) %>%
    group_by(state) %>%
    mutate(won_state = (vote_perc == max(vote_perc, na.rm = T)),
           is_nominee = (candidate == "Walter Mondale")) %>%
    ungroup() %>%
    mutate(event_type = ifelse(is_caucus(state), "caucus", "primary"),
           state = cleanup_state(state))
}
write_csv(tidy_1984(), "../data/primary/democrat_1984.csv")
```


## 1988 Primary

```{r}
tidy_1988 <- function() {
  html("../data/primary/wiki_1988.html") %>%
    html_table(header = T) %>%
    `[[`(1) %>%
    gather(`Michael Dukakis`,`Jesse Jackson`,`Al Gore`,
           `Dick Gephardt`,`Paul Simon`,`Gary Hart`,`Bruce Babbitt`,
           key = "candidate", value = "vote_perc") %>%
    mutate(Date = as.Date(paste0(Date, " 1988"), format = "%B %d %Y"),
           State = gsub("\\[\\d+\\]$","", State),
           vote_perc = as.numeric(gsub("%", "", vote_perc))) %>%
    rename(date = Date,
           state = State) %>%
    group_by(state) %>%
    mutate(won_state = (vote_perc == max(vote_perc, na.rm = T)),
           is_nominee = (candidate == "Michael Dukakis")) %>%
    ungroup() %>%
    mutate(event_type = ifelse(is_caucus(state), "caucus", "primary"),
           state = cleanup_state(state))
}
write_csv(tidy_1988(), "../data/primary/democrat_1988.csv")
```

## 1992 Primary

This is more a note to myself than anything else. I'm going to scrape data for the Democratic primaries out of some website I found.

The basic procedure is to transform the state results into a "state-candidate" result (with a boolean column tracking if the candidate won the state). This, apparently, is tidy data.

One quirk I found is that Idaho, Minnessota, North Dakota, Texas, and Washington all had BOTH a primary AND a caucus in 1992.

```{r}
parse_delegates <- function(contest) {
  ifelse(identical(character(0),grep("\\d+ Delegates", contest, value = T)),
  0,
  as.integer(gsub("Delegates\\s*)?$",
       "",
       gsub("^.*[(]\\d{1,3} of ", "", grep("\\d+ Delegates", contest, value = T)))))
}
parse_delegates("Iowa Caucus(49 of 49 Delegates)")
```
```{r}
parse_1992_votes <- function(s) {
  if ("" == s) {
    return(0)
  } else if (grepl("(\\d+[.]?\\d*)[%]?$", s)) {
    return(as.numeric(str_match(s, "(\\d+[.]?\\d*)[%]?$")[,2]))
  }
  return(as.numeric(str_match(s, "(\\d+[.]?\\d*)%")[,2]))
}
parse_1992_votes("24.78%(9-Delegates)") == 24.78
```

```{r}
tidy_1992 <- function() {
  html("../data/primary/wiki_1992.html") %>%
    html_table(header = T) %>%
    `[[`(1) %>%
    gather(`Bill Clinton`, `Jerry Brown`, `Paul Tsongas`, `Bob Kerrey`,
           `Tom Harkin`, `Larry Agran`, `Uncommitted`, `Charles Woods`,
           `Eugene McCarthy`, `Lyndon LaRouche`, key = "Candidate", value = "Vote") %>%
    mutate(Date = as.Date(paste0(Date, " 1992"), format = "%B %d %Y")) %>%
    rename(date = Date,
           contest = Contest,
           candidate = Candidate,
           vote_perc = Vote) %>%
    mutate(state = gsub("[(].*$", "", contest)) %>%
    group_by(state,candidate) %>%
    mutate(vote_perc = parse_1992_votes(vote_perc)) %>%
    ungroup() %>%
    group_by(state) %>%
    transmute(candidate,
              date,
              vote_perc,
              # delegates = parse_delegates(contest),
              won_state = (max(vote_perc, na.rm = T) == vote_perc),
              is_nominee = (candidate == "Bill Clinton")) %>%
    ungroup() %>%
    mutate(event_type = ifelse(is_caucus(state), "caucus", "primary"),
           state = cleanup_state(state))
}
write_csv(tidy_1992(), "../data/primary/democrat_1992.csv")
```

## 2004 Democratic Primary

```{r}
tidy_2004 <- function() {
  html("../data/primary/wiki_2004.html") %>%
    html_table(header = T) %>%
    `[[`(1) %>%
    filter(Date != '') %>%
    gather(`Carol Moseley Braun`, `Wesley Clark`, `Howard Dean`,
           `John Edwards`, `Richard Gephardt`, `John Kerry`, `Dennis Kucinich`,
           `Joseph Lieberman`, `Al Sharpton`,
           key = "Candidate", value = "Vote") %>%
    mutate(Date = as.Date(paste0(Date, " 2004"), format = "%B %d %Y"),
           Vote = as.numeric(gsub("%.*$","", Vote)),
           is_nominee = (Candidate == "John Kerry")) %>%
    rename(date = Date,
           state = State,
           vote_perc = Vote,
           candidate = Candidate) %>%
    group_by(state) %>%
    mutate(won_state = ifelse(is.na(vote_perc), F, (vote_perc == max(vote_perc, na.rm = T))),
           vote_perc = ifelse(is.na(vote_perc), 0, vote_perc)) %>%
    ungroup() %>%
    mutate(event_type = ifelse(is_caucus(state), "caucus", "primary"),
           state = cleanup_state(state))
}
write_csv(tidy_2004(), "../data/primary/democrat_2004.csv")
```

## 2008

```{r}
extract_2008_votes <- function(s) {
  if ("" == s) {
    return(0)
  }
  as.numeric(gsub("%", "", regmatches(s, regexpr("(\\d+)%$", s, perl = TRUE))))
}
```



```{r}
tidy_2008 <- function() {
  html("../data/primary/wiki_2008.html") %>%
    html_table(header = T) %>%
    `[[`(1) %>%
    select(-Uncommitted) %>%
    gather(`Hillary Clinton`, `Barack Obama`, `Mike Gravel`, `John Edwards`,
           `Dennis Kucinich`, `Bill Richardson`, `Joe Biden`, `Chris Dodd`,
           key = "Candidate", value = "Vote") %>%
    mutate(Date = as.Date(paste0(Date, " 2008"), format = "%B %d %Y"),
           Vote = ifelse(Vote == "", 0, extract_2008_votes(Vote)),
           is_nominee = (Candidate == "Barack Obama")) %>%
    rename(date = Date,
           state = State,
           vote_perc = Vote,
           candidate = Candidate) %>%
    group_by(state) %>%
    mutate(won_state = ifelse(is.na(vote_perc), F, (vote_perc == max(vote_perc, na.rm = T))),
           vote_perc = ifelse(is.na(vote_perc), 0, vote_perc)) %>%
    ungroup() %>%
    mutate(event_type = ifelse(is_caucus(state), "caucus", "primary"),
           state = cleanup_state(state))
} 
write_csv(tidy_2008(), "../data/primary/democrat_2008.csv")
```
