---
title: "Democratic Primary"
author: "Alex Nelson"
date: "5/18/2019"
output:
  md_document:
    variant: gfm
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(scales) # for muted()
library(rmarkdown)
knitr::opts_chunk$set(echo = TRUE)
```

# Data

```{r}
announcements <- data.frame(
  candidate = c("Elizabeth Warren","Tulsi Gabbard","Julian Castro","Kirsten Gillibrand","Kamala Harris","Pete Buttigieg","Cory Booker","Amy Klobuchar","Bernie Sanders","Jay Inslee","John Hickenlooper","Beto O'Rourke","Mike Gravel","Tim Ryan","Eric Swalwell","Seth Moulton","Joe Biden","Michael Bennet","Seth Bullock","Bill de Blasio"),
  announced = as.Date(c('2018-12-31','2019-01-11','2019-01-12','2019-01-15','2019-01-21','2019-01-23','2019-02-01','2019-02-10','2019-02-19','2019-03-01','2019-03-04','2019-03-14','2019-03-19','2019-04-04','2019-04-08','2019-04-22','2019-04-25','2019-05-02','2019-05-14','2019-05-16'))
)
```

Now we can compute the intervals.
```{r}
intervals <- announcements %>%
    arrange(announced) %>%
    mutate(diff = as.integer(announced - lag(announced, default = first(announced))))
```

```{r}
hist(as.integer(intervals$diff),xlab="Days between announcement",main="Intervals beween Democratic Primary Candidate Announcements")
```

Using ggplot2 to make things prettier.

```{r}
# "#77AB43"
diffs <- intervals$diff[2:length(intervals$diff)]
start_dt <- min(intervals$announced)
end_dt <- max(intervals$announced)
shape <- nrow(intervals)/(1.0*as.integer(end_dt - start_dt))
x <- seq(0, max(intervals$diff), length.out=100)
df <- with(intervals, data.frame(x = x, y = length(intervals$diff)*dexp(x, rate=shape)))
p <- ggplot(intervals) + geom_histogram(aes(x = diff), # bins=max(intervals$diff), 
                                   fill=muted("#008FD5", l=50, c=50), binwidth=0.75) +
  geom_line(data = df, aes(x = x, y = y), color = "#FF2700") +
  labs(caption="http://political-arithmetic.blogspot.com",
       title= "Wait until next Announcement",
       subtitle = "Intervals between Democratic candidate announcements") +
  scale_color_fivethirtyeight() +
  theme_fivethirtyeight(base_size = 8) +
  theme(axis.title = element_text()) + ylab("Number of Candidates") + xlab("Days")
p
```

Then we save the plot

```{r}
setwd("D:/src/pa/_posts/")
ggsave("../img/2019-05-primary.png", plot=p)
```

## Kalmagorov-Smirnov Test

We can test if this is a good fit for the exponential distribution (it probably isn't). We can't have duplicate values, and we're kind of short on data, so lets try "padding the data" with duplicates, and "smooth out" duplicate values by "tiny amounts".

```{r}
idx <- 1;
diff <- c();
for (x in as.integer(intervals$diff)) {
  if (x %in% diff) {
    diff[idx] <- x + (idx/10000.0);
  } else {
    diff[idx] <- x;
  }
  idx <- idx + 1;
}
for (x in as.integer(intervals$diff)) {
  if (x %in% diff) {
    tmp = 4;
    y = x + idx/(10000.0*tmp);
    while (y %in% diff) {
      tmp <- tmp + 1;
      y <- x + idx/(10000.0*tmp);
    }
    diff[idx] <- y;
  } else {
    diff[idx] <- x;
  }
  idx <- idx + 1;
}
for (x in as.integer(intervals$diff)) {
  if (x %in% diff) {
    tmp = 4;
    y = x + idx/(10000000.0*tmp);
    while (y %in% diff) {
      tmp <- tmp + 1;
      y <- x + idx/(10000000.0*tmp);
    }
    diff[idx] <- y;
  } else {
    diff[idx] <- x;
  }
  idx <- idx + 1;
}
ks.test(diff,"pexp",shape, exact=TRUE)
```

### Testing without padding

Lets see what the data tells us, without duplicating data, but padding by a "tiny amount" to avoid duplicates.

```{r}
idx <- 1;
diff <- c();
for (x in as.integer(intervals$diff[2:length(intervals$diff)])) {
  if (x %in% diff) {
    y <- x;
    while (y %in% diff) {
      y <- y + (1e-13)*idx;
    }
    diff[idx] <- y;
  } else {
    diff[idx] <- x;
  }
  idx <- idx + 1;
}
```

```{r}
ks.test(diff,"pexp",shape, exact=TRUE)
```

## Simulating Candidate Announcements

We could simply simulate the total number of candidates announced, given some "number of days into the primary" `t`, and the `rate` at which candidates announce.

```{r}
# "#77AB43"
simulate_candidates <- function(t, rate) {
    
    path <- matrix(0, nrow = 1, ncol = 2)
    
    jumps_number <- rpois(1, lambda = rate * t)
    jumps_time <- runif(n = jumps_number, min = 0, max = t) %>% sort()
    
    for(j in seq_along(jumps_time)) {
        jump <- matrix(c(jumps_time[j], path[nrow(path), 2],
                         jumps_time[j], path[nrow(path), 2]  + 1),
                       nrow = 2, ncol = 2, byrow = TRUE)
        path <- rbind(path, jump)
    }
    
    path <- rbind(path,
                  c(t, path[nrow(path), 2]))
    
    list(path, jumps_time)
    
}
path2 <- simulate_candidates(as.integer(end_dt - start_dt), shape)
```

Then we plot the intervals using the handy `diff()` function, which should resemble what we witness.

```{r}
data.frame(it = diff(path2[[2]])) %>%
    ggplot() +
    geom_histogram(aes(it, y = ..density..), fill="#008FD5") +
    stat_function(fun = dexp, args=list(rate=shape)) +
    labs(caption="http://political-arithmetic.blogspot.com",
       title= "Simulated Candidate Announcements",
       subtitle = "Intervals between Democratic candidate announcements") +
    scale_color_fivethirtyeight() +
    theme_fivethirtyeight(base_size = 8) +
    theme(axis.title = element_text()) + ylab("Proportion of Candidates") + xlab("Days")
```