---
title: "Swingiest Counties in 2016"
author: "Alex Nelson"
date: "5/13/2019"
output:
  md_document:
    variant: gfm
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)

library(dplyr)
library(car) # for vif()
library(usmap)
library(urbnmapr)

library(arm) # for glmer

library(tidycensus)
library(tidyverse)
library(ggplot2)
library(scales) # for muted()
library(rmarkdown)
knitr::opts_chunk$set(echo = TRUE)
```

# Loading our data

We need a few helper functions to denormalize useful data in the rows.

```{r}
winner_by_state_and_year <- function(results) {
  return(results %>% group_by(state,year) %>% filter(votepercentage==max(votepercentage)) %>%  arrange(party));
}
```

We need to normalize the parties, since the Democratic party in Minnessota calls themselves the Democratic-Farmer-Labor party.

```{r}
make_party_into_factor <- function(results) {
  results$party[results$party == 'democratic-farmer-labor'] <- 'democrat';
  results$party[is.na(results$party)] <- '$third-party';
  results$party <- as.factor(results$party);
  return(results);
}
add_vote_percentage <- function(results) {
  results$votepercentage <- with(results, candidatevotes/totalvotes);
  return(results);
}
```

We need a helper function to load `RData` files into specified variables.

```{r}
load_obj <- function(path) {
  env <- new.env();
  nm <- load(path,env)[1];
  env[[nm]];
}
```

```{r}
count_alternates <- function(column) {
  count <- 0;
  if (length(column) > 0) {
    val <- column[[1]][1];
    for (i in 2:length(column)) {
      if (column[[i]][1] != val) {
        val <- column[[i]][1];
        count <- count + 1;
      }
    }
  }
  return(count);
}
```

```{r}
number_of_swings_since <- function(results, election_year, state_name, number_of_elections=7) {
  start_year <- election_year - (number_of_elections - 1)*4;
  has_data <- (1976 <= start_year);
  is_election_year <- ((election_year - 1976)%%4 == 0);
  if (!has_data || !is_election_year) {
    return(0);
  } else {
    parties<-results[which(results$state==state_name & start_year <= results$year & results$year <= election_year),c("party")];
    return(count_alternates(parties[["party"]]));
  }
}

has_swung_since <- function(results, election_year, state_name, number_of_elections=7) {
  return(number_of_swings_since(results, election_year, state_name, number_of_elections) > 1);
}
```

```{r}
state_path <- "../data/elections/presidential/state/1976_2016_president.RData"
state_results <- load_obj(state_path);
state_results <- make_party_into_factor(state_results);
state_results <- add_vote_percentage(state_results);
state_winner_data <- winner_by_state_and_year(state_results);
state_winner_data <- state_winner_data[with(state_winner_data,order(state,year)), ]

swing_data <- state_winner_data[which(state_winner_data$year >= 2000),c("state","year")]
```

```{r}
for (i in 1:nrow(swing_data)) {
  row <- swing_data[i,];
  year <- row[['year']];
  state <- row[['state']];
  swing_data[which(swing_data$state == state & swing_data$year == year),c('swings')] <- number_of_swings_since(state_winner_data,year,state,2);
}
```

The states we are interested in examining are the ones which have swung from Obama to Trump.

```{r}
kable(swing_data[which(swing_data$year==2016 & swing_data$swings>=1),])
```

```{r}
path <- "../data/elections/presidential/county/countypres_2000-2016.RData"
results <- load_obj(path);
results <- make_party_into_factor(results);
results <- add_vote_percentage(results);
county_results <- results; # for later usage
winner_data <- winner_by_state_and_year(results);
winner_data <- winner_data[with(winner_data,order(state,county,year)), ]
```

We now want to drill down on the states which were won by Obama but switched to Trump.

```{r}
states <- swing_data[which(swing_data$year==2016 & swing_data$swings>=1),c('state')]$state
relevant_data <- results[results$state %in% states,]
```

We want to find, among these states, which counties were won by Obama but switched to Trump. We don't know if comparing 2008 to 2016 is more meaningful than 2012 to 2016, so we gather all the data we can.

```{r}
counties_obama_won_2008 <- relevant_data %>% filter(year==2008) %>% group_by(state,county) %>% filter(votepercentage==max(votepercentage)) %>%  arrange(party) %>% filter(party=='democrat')
counties_obama_won_2008 <- add_vote_percentage(counties_obama_won_2008)

counties_obama_won_2012 <- relevant_data %>% filter(year==2012) %>% group_by(state,county) %>% filter(votepercentage==max(votepercentage)) %>%  arrange(party) %>% filter(party=='democrat')
counties_obama_won_2012 <- add_vote_percentage(counties_obama_won_2012)

counties_trump_won <- relevant_data %>% filter(year==2016) %>% group_by(state,county) %>% filter(votepercentage==max(votepercentage)) %>%  arrange(party) %>% filter(party=='republican')
counties_trump_won <- add_vote_percentage(counties_trump_won)
```

Now we are curious about the counties which swung from Obama to Trump, which we will try measuring in terms of percent of votes cast.

```{r}
trump_intersect_2008 <- counties_trump_won[counties_trump_won$state %in% counties_obama_won_2008$state & counties_trump_won$county %in% counties_obama_won_2008$county,]
obama_intersect_2008 <- counties_obama_won_2008[counties_obama_won_2008$state %in% counties_trump_won$state & counties_obama_won_2008$county %in% counties_trump_won$county,]

compare_2008 <- merge(obama_intersect_2008, trump_intersect_2008, by=c('state','county','FIPS'))
compare_2008$trump_gain <- compare_2008$votepercentage.y - compare_2008$votepercentage.x

data_2008 <- compare_2008[which(compare_2008$trump_gain > 0),c('state','county','FIPS','trump_gain')]
```

```{r}
kable(data_2008[order(-data_2008$trump_gain),])
```
# Plots for the 2008 Results

Now we can try to plot the data:

```{r}
data_2008 <- compare_2008[,c('state','county','FIPS','trump_gain')]
county_2008 <- data_2008
county_2008$county_fips = as.character(county_2008$FIPS)
county_2008 <- left_join(counties, county_2008, by = "county_fips") 

county_2008 %>% 
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = NA) +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = 'white',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

We plot each state independently

## Florida

```{r}
county_2008 %>% filter(state_name == 'Florida') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```


```{r}
fl_2008 <- (data_2008 %>% filter(state == 'Florida'))
kable(fl_2008[order(-fl_2008$trump_gain),c('county','trump_gain')])
```

## Iowa

```{r}
county_2008 %>% filter(state_name == 'Iowa') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
ia_2008 <- (data_2008 %>% filter(state == 'Iowa'))
kable(ia_2008[order(-ia_2008$trump_gain),c('county','trump_gain')])
```

## Michigan


```{r}
county_2008 %>% filter(state_name == 'Michigan') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
mi_2008 <- (data_2008 %>% filter(state == 'Michigan'))
kable(mi_2008[order(-mi_2008$trump_gain),c('county','trump_gain')])
```

## Nebraska

```{r}
county_2008 %>% filter(state_name == 'Nebraska') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
ne_2008 <- (data_2008 %>% filter(state == 'Nebraska'))
kable(ne_2008[order(-ne_2008$trump_gain),c('county','trump_gain')])
```

## Ohio

```{r}
county_2008 %>% filter(state_name == 'Ohio') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
oh_2008 <- (data_2008 %>% filter(state == 'Ohio'))
kable(oh_2008[order(-oh_2008$trump_gain),c('county','trump_gain')])
```

## Pennsylvania

```{r}
county_2008 %>% filter(state_name == 'Pennsylvania') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
pa_2008 <- (data_2008 %>% filter(state == 'Pennsylvania'))
kable(pa_2008[order(-pa_2008$trump_gain),c('county','trump_gain')])
```

## Wisconsin


```{r}
county_2008 %>% filter(state_name == 'Wisconsin') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
wi_2008 <- (data_2008 %>% filter(state == 'Wisconsin'))
kable(wi_2008[order(-wi_2008$trump_gain),c('county','trump_gain')])
```

# Plots for the 2012 Results

And the comparison for 2012 results


```{r}
trump_intersect_2012 <- counties_trump_won[counties_trump_won$state %in% counties_obama_won_2012$state & counties_trump_won$county %in% counties_obama_won_2012$county,]
obama_intersect_2012 <- counties_obama_won_2012[counties_obama_won_2012$state %in% counties_trump_won$state & counties_obama_won_2012$county %in% counties_trump_won$county,]

compare_2012 <- merge(obama_intersect_2012, trump_intersect_2012, by=c('state','county','FIPS'))
compare_2012$trump_gain <- compare_2012$votepercentage.y - compare_2012$votepercentage.x

data_2012 <- compare_2012[which(compare_2012$trump_gain > 0),c('state','county','FIPS','trump_gain')]
```

There were `nrow(data_2012)` counties which flipped from Obama 2012 to Trump in 2016:

```{r}
kable(data_2012[order(-data_2012$trump_gain),])
```

Now we can try to plot the data:

```{r}
data_2012 <- compare_2012[,c('state','county','FIPS','trump_gain')]
county_2012 <- data_2012
county_2012$county_fips = as.character(county_2012$FIPS)
county_2012 <- left_join(counties, county_2012, by = "county_fips")

county_2012 %>% 
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = NA) +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

## Florida

```{r}
county_2012 %>% filter(state_name == 'Florida') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```


```{r}
fl_2012 <- (data_2012 %>% filter(state == 'Florida'))
kable(fl_2012[order(-fl_2012$trump_gain),c('county','trump_gain')])
```


## Iowa

```{r}
county_2012 %>% filter(state_name == 'Iowa') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
ia_2012 <- (data_2012 %>% filter(state == 'Iowa'))
kable(ia_2012[order(-ia_2012$trump_gain),c('county','trump_gain')])
```

## Michigan

```{r}
county_2012 %>% filter(state_name == 'Michigan') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
mi_2012 <- (data_2012 %>% filter(state == 'Michigan'))
kable(mi_2012[order(-mi_2012$trump_gain),c('county','trump_gain')])
```

## Ohio

```{r}
county_2012 %>% filter(state_name == 'Ohio') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
oh_2012 <- (data_2012 %>% filter(state == 'Ohio'))
kable(oh_2012[order(-oh_2012$trump_gain),c('county','trump_gain')])
```

## Pennsylvania

```{r}
county_2012 %>% filter(state_name == 'Pennsylvania') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
pa_2012 <- (data_2012 %>% filter(state == 'Pennsylvania'))
kable(pa_2012[order(-pa_2012$trump_gain),c('county','trump_gain')])
```

## Wisconsin

```{r}
county_2012 %>% filter(state_name == 'Wisconsin') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
wi_2012 <- (data_2012 %>% filter(state == 'Wisconsin'))
kable(wi_2012[order(-wi_2012$trump_gain),c('county','trump_gain')])
```

# Critical Result: Do these 2012 Obama-Trump counties matter? (Yes, they swung the election)

Do these counties really matter? If they did, they "swung" the election for Trump. Contrast this with a "surge" of voters in safe counties. Lets see if there are sufficient votes from these Obama-Trump counties to swing the election. First we need to get Clinton's results for the Obama-Trump counties:

```{r}
clinton_2016 <- relevant_data[relevant_data$state %in% counties_trump_won$state & relevant_data$county %in% counties_trump_won$county & relevant_data$party == 'democrat' & relevant_data$year == 2016,]
```

Now we merge the two datasets together to compute the margins in the Obama-Trump counties.

```{r}
merged_2016 <- merge(trump_intersect_2012, clinton_2016, by=c('state','county','FIPS'));
merged_2016$margin <- merged_2016$candidatevotes.x - merged_2016$candidatevotes.y

margin_by_counties_2016 <- merged_2016 %>% group_by(state) %>% summarize(margin = sum(margin))

margin_by_counties_2016$pivot <- 1 + (margin_by_counties_2016$margin %/% 2)

kable(margin_by_counties_2016)
```

We need to compare the results for the 7 states.

```{r}
state_2016 <- state_results[which(state_results$year == 2016 & state_results$state %in% merged_2016$state & (state_results$party == 'democrat' | state_results$party == 'republican')),]
```

We should also track the third party votes, since presumably they also acted as a [spoiler](https://en.wikipedia.org/wiki/Spoiler_effect).

```{r}
margin_2016_by_state <- state_2016 %>% group_by(state) %>% summarize(margin = max(candidatevotes) - min(candidatevotes), third_party_votes = max(totalvotes) - sum(candidatevotes))
kable(margin_2016_by_state)
```


Now we can answer the question, finally!

```{r}
pivots_by_state <- merge(margin_2016_by_state,margin_by_counties_2016,c('state'))
pivots_by_state$swingable <- pivots_by_state$margin.x < pivots_by_state$pivot
colnames(pivots_by_state)[colnames(pivots_by_state)=="margin.x"] <- "state_margin"
colnames(pivots_by_state)[colnames(pivots_by_state)=="margin.y"] <- "swing_counties_margin"
colnames(pivots_by_state)[colnames(pivots_by_state)=="third_party_votes.x"] <- "state_third_party_votes"
colnames(pivots_by_state)[colnames(pivots_by_state)=="third_party_votes.y"] <- "swing_counties_third_party_votes"
kable(pivots_by_state)
```
So, Iowa, Nebraska, and Ohio were not swingable. But if Florida, Michigan, Pennsylvania, and Wisconsin had swung for Clinton, that is a difference of 29, 16, 20, and 10 electoral votes for a sum of 75 electoral votes. Had this happen, Trump would have received 229 electoral votes to Clinton's 302.

On the other hand, the third party votes do not seem to credibly act as a spoiler.

## Florida Suspicions

There were two counties with Hacking suspicions in Florida. At first, Leon seems like a candidate, except it's the most highly educated county in the country (and there is a strong negative correlation between education and voting for Trump).

Madison county seems incredibly suspicious, it's approximately 40% African American and 60% white. Just from the Z-transform of vote percentage, this seems quite an amazing shift in votes for Clinton decreasing more than Trump's increase. (Break this out into its own post. Right now, just scratchwork.)

```{r}
fl_counties_pre_2016 <- county_results %>% filter(year < 2016 & state=='Florida')
fl_mean_sd <- fl_counties_pre_2016 %>% group_by(county,party) %>% summarize(mean=mean(votepercentage),sd=sd(votepercentage)) %>% filter(party=='democrat' | party=='republican')
fl_counties_2016 <- county_results %>% filter(year == 2016 & state=='Florida')
merged_fl <- merge(fl_counties_2016, fl_mean_sd, by=c('county','party'))
merged_fl$z <- (merged_fl$votepercentage - merged_fl$mean)/merged_fl$sd

kable(merged_fl[order(merged_fl$z),c('county','z','party')])
```

Well, what if this was a surge election, and it brought a lot of people out of the woodwork to vote for Trump? We can take this into account, normalizing statewide the county-'z' value to measure how variable each county was relative to the state as a whole:

```{r}
merged_fl$normed_z <- (merged_fl$z - mean(merged_fl$z))/sd(merged_fl$z)
kable(merged_fl[order(merged_fl$normed_z),c('county','normed_z','party')])
```

Remember: Leon county is the most educated in the country, it is not anomalous in defecting from Trump. The outliers have $|z|\geq 2$, so specifically Madison, DeSoto, and Glades counties appear statistically anomalous.

# How did Trump Win the White House? (Other Reasons)

Although we just found, by accident, one way to explain how Trump won the White House (namely, Obama voters from 2012 who switched to vote for Trump in 2016 explain how a sufficient number of swing states...swung...), are there other ways to explain _how Trump won the White House?_

We will use a post-stratified multilevel model training a logistic regression to match against exit poll data, then using those coefficients in a linear regression to try to match the voting results. This is a partially pooled model, so although the coefficients vary state-to-state, they are still "in the same ballpark". For reviews of this approach, see:

- the MRP Primer http://www.princeton.edu/~jkastell/MRP_primer/mrp_primer.pdf
- David Park, Andrew Gelman, Joseph Bafumi, [Bayesian Multilevel Estimation with Poststratification: State-Level Estimates from National Polls](http://www.stat.columbia.edu/~gelman/research/published/parkgelmanbafumi.pdf)

## Exit Poll Data

We load CNN's exit poll data. Fortunately, it is stored as a CSV file.

```{r}
exit_poll_path <- "../data/elections/presidential/exit_polls/cnn_04022017.csv"
exit_poll_df <- read.csv(file=exit_poll_path, header=TRUE, sep=",")
```

We combine race and gender into a single variable with 7 categories (the 3 race categories [white, black, hispanic] and 2 genders [men, women] combinations, and 1 `other` category). Asians constitute about 4% of the respondents, and "others" constitutes 3%. Whites constitute 71% of respondents, non-whites constitute 29%.

The age brackets are: 18-29, 30-44, 45-64, and 65+. So, 5 categories.

The education brackets are: "High school or less", "Some college", "College graduate", "Postgraduate".

The income brackets are `30k-50k`, `50k-100k`, `100k-150k`, `150k-200k`, `200k-250k`, and `250k+`. We can use the `B19037` table from the ACS5 census data to get the estimates of number of householder (i.e., occupant who either owns or rents the residency as determined by whose name is on the deed, renal agreement, etc.) by age, income, and race. The ACS5 data has finer categories of income brackets. Unfortunately, the age-brackets for this are younger than 25, 25-64, 45-64, 65+, so we need to be careful not to toss out all the census data too quickly.

### GOAL: Infer Logistic Regression

We want to construct a logistic regression

```
prob_clinton(ethnicity,gender,education,income) = invlogit(b0 + b1*ethnicity + b2*gender + b3*education + b4*income)
```

...and similarly for Trump, but we need to compute the coefficients `b0`, ..., `b4`. The constant coefficient is just the `logit(clinton_votes/(clinton_votes + trump_votes)) = logit(65853514.0/(62984828.0 + 65853514.0)) = 0.04453892`.

```{r}
beta0 <- log(65853514.0/62984828.0)
```

Although left unstated, exit polls are usually plus-or-minus 4%, and assuming it is with 95% confidence, (so approximating the numerator and denominator as normally distributed random variables with expected values given, and standard deviation being 2%) the [propagation  of uncertainty](https://en.wikipedia.org/wiki/Propagation_of_uncertainty#Example_formulae) implies the variance of `beta0` is

```{r}
var_beta <- function(A,B) {
  return( ((3.0/A)**2)+((3.0/B)**2) );
}
var_beta0 <- var_beta(65853514.0, 62984828.0) # nearly machine epsilon
```

The variance for other betas will be closer to half a percent. (It is bounded from above for `A=B=50`, `var_beta(50,50)~0.06**2`.) Even if one were to insist on the margin of error being twice that of a normal poll (i.e., it should be 6% instead of our 4%), the variance's upper bound is moved to approximately `0.08485**2`. (["This makes the margins for error somewhere between 50-90% higher than they would be for comparable telephone surveys."](https://fivethirtyeight.com/features/ten-reasons-why-you-should-ignore-exit/))

We can estimate the coefficients for any group by taking `beta_clinton(group) = log(percent_clinton(group)/percent_trump(group)) - beta0`. This is the basic strategy of one-factor models, after all.

```{r}
national_exit_polls <- exit_poll_df[which(exit_poll_df$state == 'nation'),]

race_and_gender <- national_exit_polls[which(national_exit_polls$questions=="Race and gender"),]
age <- national_exit_polls[which(national_exit_polls$questions_id==3),]
income <- national_exit_polls[which(national_exit_polls$questions_id==13),]
married <- national_exit_polls[which(national_exit_polls$questions_id==20),]
religion <- national_exit_polls[which(national_exit_polls$questions_id==23),]
education <- national_exit_polls[which(national_exit_polls$questions == "Education"),]
```

We can estimate the error using the Wilson confidence interval.

```{r}
wilson_center <- function(p, n, z) {
  (p + 0.5*z*z/n)/(1.0 + z*z*1.0/n);
}
wilson_width <- function(p, n, z) {
  (z/(1.0 + z*z*1.0/n))*sqrt(p*(1.0-p)/n  + (z*0.5/n)**2);
}
log_coefs <- function(r) {
  c <- wilson_center(r$Clinton_perc*0.01, r$options_perc*0.01*r$num_respondents,1)
  t <- wilson_center(r$Trump_perc*0.01, r$options_perc*0.01*r$num_respondents,1)
  r$clinton_numerator <- c/t
  r$trump_numerator <- t/c
  
  r$clinton_coef <- log(r$clinton_numerator) - beta0
  r$trump_coef <- log(r$trump_numerator) - beta0
  r$log_var <- (wilson_width(r$Clinton_perc*0.01, r$options_perc*0.01*r$num_respondents,1)/(0.01*r$Clinton_perc))**2 + (wilson_width(r$Trump_perc*0.01, r$options_perc*0.01*r$num_respondents,1)/(0.01*r$Trump_perc))**2;
  r$log_sd <- sqrt(r$log_var);
  return(r);
}
```

#### Religion

#### Married

```{r}
education <- log_coefs(married)

kable(married[,c("options","clinton_coef","trump_coef")])
```

#### Gender and Ethnicity

Then we have the numerators for the coefficients

```{r}
race_and_gender <- log_coefs(race_and_gender)

kable(race_and_gender[,c("options","clinton_coef","trump_coef")])
```


#### Education

For education we have:

```{r}
education <- log_coefs(education)

kable(education[,c("options","clinton_coef","trump_coef")])
```

Observe that education is linear progressing. We can assign the following weights to education level: -1 to "High school or less", 0 to "Some college", 1 to "College graduate", and 2 to "Postgraduate". Then we find a linear regression the coefficient is better approximated by `x1` in the following table:

```{r}
df <- data.frame(y=education$clinton_coef, z=education$trump_coef,x=1:4,x0=0:3,x1=-1:2)

clm <- lm(y~x1,df)
tlm <- lm(z~x1,df)

beta_education <- clm$coefficients[2] # == -tlm$coefficients[2]

kable(data.frame(Clinton_coefficients=clm$coefficients, Trump_coefficient=tlm$coefficients))
```
#### Income


```{r}
income <- log_coefs(income)

kable(income[,c("options","clinton_coef","trump_coef")])
```
```{r}
df <- data.frame(y = income$clinton_coef, z=income$trump_coef, x=1:6, x2=c(15,0.5*(30+45),0.5*(50+100),0.5*(100+200),0.5*(200+250),300),x3=c(15,0.5*(30+45),0.5*(50+100),0.5*(100+200),0.5*(200+250),300)**2)

clm <- lm(y~x**2,df)
```

#### Age

For age we have a partition of ages into 5 or 6 classes. We take the midpoint of these intervals to approximate the value, and take a linear regression of the logistic coefficients.

```{r}
age <- log_coefs(age)

kable(age[,c("options","clinton_coef","trump_coef")])
```
We try to approximate this using a linear regression, taking `age-46` as the input signal appears to be best (note the intercept is approximately `-beta0`):

```{r}
df <- data.frame(y=age$clinton_coef, z=age$trump_coef,x=c(21,27,34.5,44.5,57,75),x0=-3:2,x1=c(21,27,34.5,44.5,57,75)-46)
df2 <- data.frame(y=age$clinton_coef[1:5], z=age$trump_coef[1:5], x=c(21,27,34.5,44.5,57),x0=-3:1,x1=c(21,27,34.5,44.5,57)-46)
clm <- lm(y~x1,df2)
tlm <- lm(z~x1,df2)

beta_age <- clm$coefficients[2] # == -tlm$coefficients[2] to machine epsilon

kable(data.frame(Clinton_coefficients=clm$coefficients, Trump_coefficient=tlm$coefficients))
```


#### Different States

We can consider coefficients for different states, lets see Florida:

```{r}
fl_exit_polls <- exit_poll_df[which(exit_poll_df$state == 'Florida'),]

fl_race_and_gender <- fl_exit_polls[which(fl_exit_polls$questions=="Race and gender"),]

fl_race_and_gender <- log_coefs(fl_race_and_gender)

kable(fl_race_and_gender[,c("options","clinton_coef","trump_coef")])
```



## Census Data

We get basic data from the census [ACS5](https://api.census.gov/data/2017/acs/acs5/variables.html)
specifically the breakdown by age and demographics. I borrow their terminology in naming the variables.

```{r}
api_key <- read.table("../data/keys/census.txt", header = FALSE)[[1]]
census_api_key(api_key)
```

There are a variety of different "types" of variables we could use from the US Census, they are called [subjects](https://www.census.gov/programs-surveys/acs/guidance/subjects.html).

Note, the estimates are given with 90% margin of error, meaning we are 90% confident the "true value" lies within the "margin of error" of the estimate. We can convert the margin of error into a standard deviation by `sd = moe/qnorm(1 - 0.5*(1 - confidence_level))` where `confidence_level=0.9` (which can be changed by the user). Ostensibly, each variable should be represented by a normal random variable with expected value given by the estimate and the sigma given by the `sd` formula. (This is all "roughly true", there are some nuances surrounding this.)

We also need to request data which will correspond with possible options in the exit polls.

```{r}
census.fl <- get_acs(geography = "county", 
              variables = c(male.18_19 = "B01001_007",
                            male.20 = "B01001_008",
                            male.21 = "B01001_009",
                            male.22_24 = "B01001_010",
                            male.25_29 = "B01001_011",
                            male.30_34 = "B01001_012",
                            male.35_39 = "B01001_013",
                            male.40_44 = "B01001_014",
                            male.45_49 = "B01001_015",
                            male.50_54 = "B01001_016",
                            male.55_59 = "B01001_017",
                            male.60_61 = "B01001_018",
                            male.62_64 = "B01001_019",
                            male.65_66 = "B01001_020",
                            male.67_69 = "B01001_021",
                            male.70_74 = "B01001_022",
                            male.75_79 = "B01001_023",
                            male.80_84 = "B01001_024",
                            male.85_plus = "B01001_025",
                            female.18_19 = "B01001_031",
                            female.20 = "B01001_032",
                            female.21 = "B01001_033",
                            female.22_24 = "B01001_034",
                            female.25_29 = "B01001_035",
                            female.30_34 = "B01001_036",
                            female.35_39 = "B01001_037",
                            female.40_44 = "B01001_038",
                            female.45_49 = "B01001_039",
                            female.50_54 = "B01001_040",
                            female.55_59 = "B01001_041",
                            female.60_61 = "B01001_042",
                            female.62_64 = "B01001_043",
                            female.65_66 = "B01001_044",
                            female.67_69 = "B01001_045",
                            female.70_74 = "B01001_046",
                            female.75_79 = "B01001_047",
                            female.80_84 = "B01001_048",
                            female.85_plus = "B01001_049",
                            male.white.18_19 = "B01001A_007",
                            male.white.20_24 = "B01001A_008",
                            male.white.25_29 = "B01001A_009",
                            male.white.30_34 = "B01001A_010",
                            male.white.35_44 = "B01001A_011",
                            male.white.45_54 = "B01001A_012",
                            male.white.55_64 = "B01001A_013",
                            male.white.65_74 = "B01001A_014",
                            male.white.75_84 = "B01001A_015",
                            male.white.85_plus = "B01001A_016",
                            female.white.18_19 = "B01001A_022",
                            female.white.20_24 = "B01001A_023",
                            female.white.25_29 = "B01001A_024",
                            female.white.30_34 = "B01001A_025",
                            female.white.35_44 = "B01001A_026",
                            female.white.45_54 = "B01001A_027",
                            female.white.55_64 = "B01001A_028",
                            female.white.65_74 = "B01001A_029",
                            female.white.75_84 = "B01001A_030",
                            female.white.85_plus = "B01001A_031",
                            male.black.18_19 = "B01001B_007",
                            male.black.20_24 = "B01001B_008",
                            male.black.25_29 = "B01001B_009",
                            male.black.30_34 = "B01001B_010",
                            male.black.35_44 = "B01001B_011",
                            male.black.45_54 = "B01001B_012",
                            male.black.55_64 = "B01001B_013",
                            male.black.65_74 = "B01001B_014",
                            male.black.75_84 = "B01001B_015",
                            male.black.85_plus = "B01001B_016",
                            female.black.18_19 = "B01001B_022",
                            female.black.20_24 = "B01001B_023",
                            female.black.25_29 = "B01001B_024",
                            female.black.30_34 = "B01001B_025",
                            female.black.35_44 = "B01001B_026",
                            female.black.45_54 = "B01001B_027",
                            female.black.55_64 = "B01001B_028",
                            female.black.65_74 = "B01001B_029",
                            female.black.75_84 = "B01001B_030",
                            female.black.85_plus = "B01001B_031",
                            male.american_indian.18_19 = "B01001C_007",
                            male.american_indian.20_24 = "B01001C_008",
                            male.american_indian.25_29 = "B01001C_009",
                            male.american_indian.30_34 = "B01001C_010",
                            male.american_indian.35_44 = "B01001C_011",
                            male.american_indian.45_54 = "B01001C_012",
                            male.american_indian.55_64 = "B01001C_013",
                            male.american_indian.65_74 = "B01001C_014",
                            male.american_indian.75_84 = "B01001C_015",
                            male.american_indian.85_plus = "B01001C_016",
                            female.american_indian.18_19 = "B01001C_022",
                            female.american_indian.20_24 = "B01001C_023",
                            female.american_indian.25_29 = "B01001C_024",
                            female.american_indian.30_34 = "B01001C_025",
                            female.american_indian.35_44 = "B01001C_026",
                            female.american_indian.45_54 = "B01001C_027",
                            female.american_indian.55_64 = "B01001C_028",
                            female.american_indian.65_74 = "B01001C_029",
                            female.american_indian.75_84 = "B01001C_030",
                            female.american_indian.85_plus = "B01001C_031",
                            male.asian.18_19 = "B01001D_007",
                            male.asian.20_24 = "B01001D_008",
                            male.asian.25_29 = "B01001D_009",
                            male.asian.30_34 = "B01001D_010",
                            male.asian.35_44 = "B01001D_011",
                            male.asian.45_54 = "B01001D_012",
                            male.asian.55_64 = "B01001D_013",
                            male.asian.65_74 = "B01001D_014",
                            male.asian.75_84 = "B01001D_015",
                            male.asian.85_plus = "B01001D_016",
                            female.asian.18_19 = "B01001D_022",
                            female.asian.20_24 = "B01001D_023",
                            female.asian.25_29 = "B01001D_024",
                            female.asian.30_34 = "B01001D_025",
                            female.asian.35_44 = "B01001D_026",
                            female.asian.45_54 = "B01001D_027",
                            female.asian.55_64 = "B01001D_028",
                            female.asian.65_74 = "B01001D_029",
                            female.asian.75_84 = "B01001D_030",
                            female.asian.85_plus = "B01001D_031",
                            male.pacific_islander.18_19 = "B01001E_007",
                            male.pacific_islander.20_24 = "B01001E_008",
                            male.pacific_islander.25_29 = "B01001E_009",
                            male.pacific_islander.30_34 = "B01001E_010",
                            male.pacific_islander.35_44 = "B01001E_011",
                            male.pacific_islander.45_54 = "B01001E_012",
                            male.pacific_islander.55_64 = "B01001E_013",
                            male.pacific_islander.65_74 = "B01001E_014",
                            male.pacific_islander.75_84 = "B01001E_015",
                            male.pacific_islander.85_plus = "B01001E_016",
                            female.pacific_islander.18_19 = "B01001E_022",
                            female.pacific_islander.20_24 = "B01001E_023",
                            female.pacific_islander.25_29 = "B01001E_024",
                            female.pacific_islander.30_34 = "B01001E_025",
                            female.pacific_islander.35_44 = "B01001E_026",
                            female.pacific_islander.45_54 = "B01001E_027",
                            female.pacific_islander.55_64 = "B01001E_028",
                            female.pacific_islander.65_74 = "B01001E_029",
                            female.pacific_islander.75_84 = "B01001E_030",
                            female.pacific_islander.85_plus = "B01001E_031",
                            male.white_only.18_19 = "B01001H_007",
                            male.white_only.20_24 = "B01001H_008",
                            male.white_only.25_29 = "B01001H_009",
                            male.white_only.30_34 = "B01001H_010",
                            male.white_only.35_44 = "B01001H_011",
                            male.white_only.45_54 = "B01001H_012",
                            male.white_only.55_64 = "B01001H_013",
                            male.white_only.65_74 = "B01001H_014",
                            male.white_only.75_84 = "B01001H_015",
                            male.white_only.85_plus = "B01001H_016",
                            female.white_only.18_19 = "B01001H_022",
                            female.white_only.20_24 = "B01001H_023",
                            female.white_only.25_29 = "B01001H_024",
                            female.white_only.30_34 = "B01001H_025",
                            female.white_only.35_44 = "B01001H_026",
                            female.white_only.45_54 = "B01001H_027",
                            female.white_only.55_64 = "B01001H_028",
                            female.white_only.65_74 = "B01001H_029",
                            female.white_only.75_84 = "B01001H_030",
                            female.white_only.85_plus = "B01001H_031",
                            male.hispanic.18_19 = "B01001I_007",
                            male.hispanic.20_24 = "B01001I_008",
                            male.hispanic.25_29 = "B01001I_009",
                            male.hispanic.30_34 = "B01001I_010",
                            male.hispanic.35_44 = "B01001I_011",
                            male.hispanic.45_54 = "B01001I_012",
                            male.hispanic.55_64 = "B01001I_013",
                            male.hispanic.65_74 = "B01001I_014",
                            male.hispanic.75_84 = "B01001I_015",
                            male.hispanic.85_plus = "B01001I_016",
                            female.hispanic.18_19 = "B01001I_022",
                            female.hispanic.20_24 = "B01001I_023",
                            female.hispanic.25_29 = "B01001I_024",
                            female.hispanic.30_34 = "B01001I_025",
                            female.hispanic.35_44 = "B01001I_026",
                            female.hispanic.45_54 = "B01001I_027",
                            female.hispanic.55_64 = "B01001I_028",
                            female.hispanic.65_74 = "B01001I_029",
                            female.hispanic.75_84 = "B01001I_030",
                            female.hispanic.85_plus = "B01001I_031"
                            ), 
              year = 2016,
              output = "wide",
              state = "FL")

census.fl.econ <- get_acs(geography = "county", 
              variables = c(income.median = "B19013_001",
                            below_poverty_line = "B06012_002E",
                            between_100_to_149_percent_poverty_line = "B06012_003",
                            above_150_percent_poverty_line = "B06012_004",
                            # income.age-bracket.income-bracket
                            income.under_25.under_10k = "B19037_003",
                            income.under_25.10k_15k = "B19037_004",
                            income.under_25.15k_20k = "B19037_005",
                            income.under_25.20k_25k = "B19037_006",
                            income.under_25.25k_30k = "B19037_007",
                            income.under_25.30k_35k = "B19037_008",
                            income.under_25.35k_40k = "B19037_009",
                            income.under_25.40k_45k = "B19037_010",
                            income.under_25.45k_50k = "B19037_011",
                            income.under_25.50k_60k = "B19037_012",
                            income.under_25.60k_75k = "B19037_013",
                            income.under_25.75k_100k = "B19037_014",
                            income.under_25.100k_125k = "B19037_015",
                            income.under_25.125k_150k = "B19037_016",
                            income.under_25.150k_200k = "B19037_017",
                            income.under_25.200k_plus = "B19037_018",
                            income.25_44.under_10k = "B19037_020",
                            income.25_44.10k_15k = "B19037_021",
                            income.25_44.15k_20k = "B19037_022",
                            income.25_44.20k_25k = "B19037_023",
                            income.25_44.25k_30k = "B19037_024",
                            income.25_44.30k_35k = "B19037_025",
                            income.25_44.35k_40k = "B19037_026",
                            income.25_44.40k_45k = "B19037_027",
                            income.25_44.45k_50k = "B19037_028",
                            income.25_44.50k_60k = "B19037_029",
                            income.25_44.60k_75k = "B19037_030",
                            income.25_44.75k_100k = "B19037_031",
                            income.25_44.100k_125k = "B19037_032",
                            income.25_44.125k_150k = "B19037_033",
                            income.25_44.150k_200k = "B19037_034",
                            income.25_44.200k_plus = "B19037_035",
                            income.45_64.under_10k = "B19037_037",
                            income.45_64.10k_15k = "B19037_038",
                            income.45_64.15k_20k = "B19037_039",
                            income.45_64.20k_25k = "B19037_040",
                            income.45_64.25k_30k = "B19037_041",
                            income.45_64.30k_35k = "B19037_042",
                            income.45_64.35k_40k = "B19037_043",
                            income.45_64.40k_45k = "B19037_044",
                            income.45_64.45k_50k = "B19037_045",
                            income.45_64.50k_60k = "B19037_046",
                            income.45_64.60k_75k = "B19037_047",
                            income.45_64.75k_100k = "B19037_048",
                            income.45_64.100k_125k = "B19037_049",
                            income.45_64.125k_150k = "B19037_050",
                            income.45_64.150k_200k = "B19037_051",
                            income.45_64.200k_plus = "B19037_052",
                            income.65_plus.under_10k = "B19037_054",
                            income.65_plus.10k_15k = "B19037_055",
                            income.65_plus.15k_20k = "B19037_056",
                            income.65_plus.20k_25k = "B19037_057",
                            income.65_plus.25k_30k = "B19037_058",
                            income.65_plus.30k_35k = "B19037_059",
                            income.65_plus.35k_40k = "B19037_060",
                            income.65_plus.40k_45k = "B19037_061",
                            income.65_plus.45k_50k = "B19037_062",
                            income.65_plus.50k_60k = "B19037_063",
                            income.65_plus.60k_75k = "B19037_064",
                            income.65_plus.75k_100k = "B19037_065",
                            income.65_plus.100k_125k = "B19037_066",
                            income.65_plus.125k_150k = "B19037_067",
                            income.65_plus.150k_200k = "B19037_068",
                            income.65_plus.200k_plus = "B19037_069"
                            ), 
              year = 2016,
              output = "wide",
              state = "FL")


census.fl.education <- get_acs(geography = "county", 
              variables = c(male.18_24.less_than_9th_grade = "B15001_004",
                            male.18_24.less_than_high_school = "B15001_005",
                            male.18_24.high_school_grad = "B15001_006",
                            male.18_24.some_college = "B15001_007",
                            male.18_24.associate_degree = "B15001_008",
                            male.18_24.bachelor_degree = "B15001_009",
                            male.18_24.graduate_degree = "B15001_010",
                            male.25_34.less_than_9th_grade = "B15001_012",
                            male.25_34.less_than_high_school = "B15001_013",
                            male.25_34.high_school_grad = "B15001_014",
                            male.25_34.some_college = "B15001_015",
                            male.25_34.associate_degree = "B15001_016",
                            male.25_34.bachelor_degree = "B15001_017",
                            male.25_34.graduate_degree = "B15001_018",
                            male.35_44.less_than_9th_grade = "B15001_020",
                            male.35_44.less_than_high_school = "B15001_021",
                            male.35_44.high_school_grad = "B15001_022",
                            male.35_44.some_college = "B15001_023",
                            male.35_44.associate_degree = "B15001_024",
                            male.35_44.bachelor_degree = "B15001_025",
                            male.35_44.graduate_degree = "B15001_026",
                            male.45_64.less_than_9th_grade = "B15001_028",
                            male.45_64.less_than_high_school = "B15001_029",
                            male.45_64.high_school_grad = "B15001_030",
                            male.45_64.some_college = "B15001_031",
                            male.45_64.associate_degree = "B15001_032",
                            male.45_64.bachelor_degree = "B15001_033",
                            male.45_64.graduate_degree = "B15001_034",
                            male.65_plus.less_than_9th_grade = "B15001_036",
                            male.65_plus.less_than_high_school = "B15001_037",
                            male.65_plus.high_school_grad = "B15001_038",
                            male.65_plus.some_college = "B15001_039",
                            male.65_plus.associate_degree = "B15001_040",
                            male.65_plus.bachelor_degree = "B15001_041",
                            male.65_plus.graduate_degree = "B15001_042",
                            # 25-64 
                            female.18_24.less_than_9th_grade = "B15001_045",
                            female.18_24.less_than_high_school = "B15001_046",
                            female.18_24.high_school_grad = "B15001_047",
                            female.18_24.some_college = "B15001_048",
                            female.18_24.associate_degree = "B15001_049",
                            female.18_24.bachelor_degree = "B15001_050",
                            female.18_24.graduate_degree = "B15001_051",
                            female.25_34.less_than_9th_grade = "B15001_053",
                            female.25_34.less_than_high_school = "B15001_054",
                            female.25_34.high_school_grad = "B15001_055",
                            female.25_34.some_college = "B15001_056",
                            female.25_34.associate_degree = "B15001_057",
                            female.25_34.bachelor_degree = "B15001_058",
                            female.25_34.graduate_degree = "B15001_059",
                            female.35_44.less_than_9th_grade = "B15001_061",
                            female.35_44.less_than_high_school = "B15001_062",
                            female.35_44.high_school_grad = "B15001_063",
                            female.35_44.some_college = "B15001_064",
                            female.35_44.associate_degree = "B15001_065",
                            female.35_44.bachelor_degree = "B15001_066",
                            female.35_44.graduate_degree = "B15001_067",
                            female.45_64.less_than_9th_grade = "B15001_069",
                            female.45_64.less_than_high_school = "B15001_070",
                            female.45_64.high_school_grad = "B15001_071",
                            female.45_64.some_college = "B15001_072",
                            female.45_64.associate_degree = "B15001_073",
                            female.45_64.bachelor_degree = "B15001_074",
                            female.45_64.graduate_degree = "B15001_075",
                            female.65_plus.less_than_9th_grade = "B15001_077",
                            female.65_plus.less_than_high_school = "B15001_078",
                            female.65_plus.high_school_grad = "B15001_079",
                            female.65_plus.some_college = "B15001_080",
                            female.65_plus.associate_degree = "B15001_081",
                            female.65_plus.bachelor_degree = "B15001_082",
                            female.65_plus.graduate_degree = "B15001_083"
                            # geographic mobility, B07001_001E et seq.
                            # geographic mobility AND income (as multiple of poverty level) B07012 et seq
                            # B08302 time leaving home to go to work
                            # B08134 how long it takes to get to work by travel time
  
                            # how to get to work B08006
                            # grandparents B10051
                            # household type [nonfamily, single male, single female, householder living alone, householder not alone]
                            # B12002 marital status
                            # B12006 marital status by labor force participation
                            # B12007 median age at first marriage (by sex, age, race)
                            # B12504 median duration of current marriage by (sex,age) by marital status
                            # B13002 women 15 to 50 who had a birth in past 12 months by marital status and age
                            # B15001 SEX BY AGE BY EDUCATIONAL ATTAINMENT FOR THE POPULATION 18 YEARS AND OVER	
                            # B16001 language spoken at home
                            # B17001 poverty status in past 12 months by sex by age
                            # B18 disability status
                            # B19 household income
                            # B20 sex by earnings in past 12 months
                            # B21 sex by veteran status
                            # B22 receipts by food stamps
                            # B23 sex by age by employment status
                            # B24 sex by occupation and median earnings in past 12 months
                            # B25 Housing units
                            # B26 group quarters population
                            # B27 health insurance coverage status
                            # B28 types of computers in household
                            # C02003 detailed race
                            # C15 sex by educational attainment
                            # C16 languages spoken at home
                            # C17 ratio of income to poverty level in past 12 months
                            # C18 age by number of disabilities
                            # C21 sex by age by veteran status for civilian population
                            ), 
              year = 2016,
              output = "wide",
              state = "FL")
```

Train on the exit polls to create the coefficients for the logistic regression, then use that to predict the turnout by county.

# Voter turnout

We are told there were 138,846,571 ballots cast in 2016 but 230,931,921	eligible voters in 2016 (and 250,055,734 people of voting age). (http://www.electproject.org/2016g)
