---
title: "Swing States"
author: "Alex Nelson"
date: "5/4/2019"
output:
  md_document:
    variant: gfm
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(dplyr)
library(car) # for vif()
library(rmarkdown)
library(arm) # for bayesglm
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Data: Winners by state

```{r}
winner_by_state_and_year <- function(results) {
  return(results %>% group_by(state,year) %>% filter(votepercentage==max(votepercentage)) %>%  arrange(party));
}
```

We need to normalize the parties, since the Democratic party in Minnessota calls themselves the Democratic-Farmer-Labor party.

```{r}
make_party_into_factor <- function(results) {
  results$party[results$party == 'democratic-farmer-labor'] <- 'democrat';
  results$party[is.na(results$party)] <- '$third-party';
  results$party <- as.factor(results$party);
  return(results);
}
add_vote_percentage <- function(results) {
  results$votepercentage <- with(results, candidatevotes/totalvotes);
  return(results);
}
```

We need a helper function to load `RData` files into specified variables.

```{r}
load_obj <- function(path) {
  env <- new.env();
  nm <- load(path,env)[1];
  env[[nm]];
}
```

```{r}
path <- "../data/elections/presidential/state/1976_2016_president.RData"
results <- load_obj(path);
results <- make_party_into_factor(results);
results <- add_vote_percentage(results);
winner_data <- winner_by_state_and_year(results);
winner_data <- winner_data[with(winner_data,order(state,year)), ]
```

```{r}
# winner_data <- winner_data[with(winner_data,order(state,year)), ]
table_data <- subset(winner_data,select=c("state","party","year","votepercentage"))
kable(table_data)
```

The number of times a column alternates between two values (or more), to help ascertain how "swingable" a state is when voting.

```{r}
count_alternates <- function(column) {
  count <- 0;
  if (length(column) > 0) {
    val <- column[[1]][1];
    for (i in 2:length(column)) {
      if (column[[i]][1] != val) {
        val <- column[[i]][1];
        count <- count + 1;
      }
    }
  }
  return(count);
}
```

```{r}
number_of_swings_since <- function(results, election_year, state_name, number_of_elections=7) {
  start_year <- election_year - (number_of_elections - 1)*4;
  has_data <- (1976 <= start_year);
  is_election_year <- ((election_year - 1976)%%4 == 0);
  if (!has_data || !is_election_year) {
    return(0);
  } else {
    parties <- results %>%
      filter(state == state_name,
             year >= start_year,
             year <= election_year) %>%
      pull(party)
#    parties<-results[which(results$state==state_name & start_year <= results$year & results$year <= election_year),c("party")];
    return(count_alternates(parties));
  }
}

weighted_number_of_swings_since <- function(results, election_year, state_name, number_of_elections=7) {
  start_year <- election_year - (number_of_elections - 1)*4;
  has_data <- (1976 <= start_year);
  is_election_year <- ((election_year - 1976)%%4 == 0);
  if (!has_data || !is_election_year) {
    return(0);
  } else {
    parties <<- results %>%
      filter(state == state_name,
             year >= start_year,
             year <= election_year) %>%
      pull(party)
    return(sum((diff(parties=="republican")!=0)/(1:(number_of_elections-1))));
  }
}

has_swung_since <- function(results, election_year, state_name, number_of_elections=7) {
  return(number_of_swings_since(results, election_year, state_name, number_of_elections) > 1);
}
```

## Margin of Victory

We need to get the runner up by year, then use that and the state's winner (by year), to compute the margin of victory.

```{r}
runner_up_by_state_and_year <- function(results) {
  results %>%
    group_by(state,year) %>%
    filter(votepercentage < max(votepercentage)) %>%
    filter(votepercentage==max(votepercentage)) %>%
    arrange(party)
}

runner_up_data = runner_up_by_state_and_year(results)
```

```{r}
aliased_data = rename(runner_up_data, runner_up_percentage = c("votepercentage"));
margin_of_victory <- merge(aliased_data,winner_data,by=c("state","year"));
margin_of_victory$percentage = margin_of_victory$votepercentage - margin_of_victory$runner_up_percentage
```

```{r}
# winner_data <- winner_data[with(winner_data,order(state,year)), ]
margin_of_victory <- margin_of_victory[with(margin_of_victory,order(state,year)), ]
table_data <- subset(margin_of_victory,select=c("state","party.y","year","percentage"))
kable(table_data)
```

## Competitive States

We could then list the states which have had margin of victory determined by less than 5% of the vote.

```{r}
kable(margin_of_victory[which(-0.05 <= margin_of_victory$percentage & margin_of_victory$percentage <= 0.05),c("party.y","year","state","percentage")])
```

# Predictions

We can try to 


```{r}
data <- margin_of_victory[,c("state","year","percentage","party.y")]
data = rename(data, margin = c("percentage"));
data = rename(data, state_winner = c("party.y"));
winner_by_year <- data.frame(
  year = c(1976,1980,1984,1988,1992,1996,2000,2004,2008,2012,2016),
  winning_party = c('democrat','republican','republican','republican','democrat','democrat','republican','republican','democrat','democrat','republican')
);
data <- merge(data,winner_by_year,by=c("year"));
data$bellwether <- with(data, ifelse(as.character(state_winner)==as.character(winning_party),1,0))
# data$swings <- with(data, number_of_swings_since(winner_data, year, state))
swing_data <- winner_data[which(winner_data$year >= 2000),c("state","year")]


for (i in 1:nrow(swing_data)) {
  row <- swing_data[i,];
  year <- row[['year']];
  state <- row[['state']];
  swing_data[which(swing_data$state == state & swing_data$year == year),c('weighted_swings')] <- weighted_number_of_swings_since(winner_data,year,state);
  swing_data[which(swing_data$state == state & swing_data$year == year),c('swings')] <- number_of_swings_since(winner_data,year,state);
}
data <- merge(data,swing_data,by=c("year","state"));
```

## Training Data

Florida and Ohion have both been a swing state since 2000.

According to [FairVote](https://www.fairvote.org/presidential-campaign-strategies-based-on-swing-states)

- '2012',	'Virginia'

But non-examples include the Mountain and Plains states (which are solidly Republican in recent times), specifically:
- Idaho, 
- Wyoming,
- the Dakotas,
- Montana,
- Utah,
- Kansas,
- Oklahoma, and
- Nebraska,

California has been solidly Democratic.

## Campaign events

In 2012, NV had 13, CO 23, FL 40, NC 3, VA 36, OH 73, PA 5, MI 1, WI 18, IA 27, MN 1, NH 13
```{r}
data$events <- 0
election <- 2012
data[which(data$state=="Nevada" & data$year==election),c("events")] <- 13
data[which(data$state=="Colorado" & data$year==election),c("events")] <- 23
data[which(data$state=="Florida" & data$year==election),c("events")] <- 40
data[which(data$state=="North Carolina" & data$year==election),c("events")] <- 3
data[which(data$state=="Virginia" & data$year==election),c("events")] <- 36
data[which(data$state=="Ohio" & data$year==election),c("events")] <- 73
data[which(data$state=="Pennsylvania" & data$year==election),c("events")] <- 5
data[which(data$state=="Michigan" & data$year==election),c("events")] <- 1
data[which(data$state=="Wisconsin" & data$year==election),c("events")] <- 18
data[which(data$state=="Iowa" & data$year==election),c("events")] <- 27
data[which(data$state=="Minnesota" & data$year==election),c("events")] <- 1
data[which(data$state=="New Hampshire" & data$year==election),c("events")] <- 13
```

In 2016, we have: CA 1, NV 17, AZ 10, UT 1, CO 19, NM 3, TX 1, MS 1, FL 71,
GA 3, NC 55, VA 23, MN 2, IA 21, MO 2, NE 2, WI 14, MI 22, IL 1, IN 2, OH 48, PA 54, NH 21, CT 1, ME 3

```{r}
election <- 2016
data[which(data$state=="California" & data$year==election),c("events")] <- 1
data[which(data$state=="Nevada" & data$year==election),c("events")] <- 17
data[which(data$state=="Arizona" & year==election),c("events")] <- 10
data[which(data$state=="Utah" & data$year==election),c("events")] <- 1
data[which(data$state=="Colorado" & data$year==election),c("events")] <- 19
data[which(data$state=="New Mexico" & data$year==election),c("events")] <- 3
data[which(data$state=="Texas" & data$year==election),c("events")] <- 1
data[which(data$state=="Mississippi" & data$year==election),c("events")] <- 1
data[which(data$state=="Florida" & data$year==election),c("events")] <- 71
data[which(data$state=="Georgia" & data$year==election),c("events")] <- 3
data[which(data$state=="North Carolina" & data$year==election),c("events")] <- 55
data[which(data$state=="Virginia" & data$year==election),c("events")] <- 23
data[which(data$state=="Minnesota" & data$year==election),c("events")] <- 2
data[which(data$state=="Iowa" & data$year==election),c("events")] <- 21
data[which(data$state=="Missouri" & data$year==election),c("events")] <- 2
data[which(data$state=="Nebraska" & data$year==election),c("events")] <- 2
data[which(data$state=="Wisconsin" & data$year==election),c("events")] <- 14
data[which(data$state=="Michigan" & data$year==election),c("events")] <- 22
data[which(data$state=="Illinois" & data$year==election),c("events")] <- 1
data[which(data$state=="Indiana" & data$year==election),c("events")] <- 2
data[which(data$state=="Ohio" & data$year==election),c("events")] <- 48
data[which(data$state=="Pennsylvania" & data$year==election),c("events")] <- 54
data[which(data$state=="New Hampshire" & data$year==election),c("events")] <- 21
data[which(data$state=="Connecticut" & data$year==election),c("events")] <- 1
data[which(data$state=="Maine" & data$year==election),c("events")] <- 3
```

In [2008](http://archive.fairvote.org/tracker/): AZ 2. CO 17, DC 1, FL 30, IL 1, IN 10, IA 8, ME 1, MD 1, MI 12, MN 1, MO 16, NV 9, NH 9, NM 9, NY 10, NC 13, OH 35, PA 29, TN 1, TX 1, VA 21, WV 1, WI 11

```{r}
election <- 2008
data[which(data$state=="Arizona" & data$year==election),c("events")] <- 2
data[which(data$state=="Colorado" & data$year==election),c("events")] <- 17
data[which(data$state=="District of Columbia" & data$year==election),c("events")] <- 1
data[which(data$state=="Florida" & data$year==election),c("events")] <- 30
data[which(data$state=="Illinois" & data$year==election),c("events")] <- 1
data[which(data$state=="Indiana" & data$year==election),c("events")] <- 10
data[which(data$state=="Iowa" & data$year==election),c("events")] <- 8
data[which(data$state=="Maine" & data$year==election),c("events")] <- 1
data[which(data$state=="Maryland" & data$year==election),c("events")] <- 1
data[which(data$state=="Michigan" & data$year==election),c("events")] <- 12
data[which(data$state=="Minnesota" & data$year==election),c("events")] <- 1
data[which(data$state=="Missouri" & data$year==election),c("events")] <- 16
data[which(data$state=="Nevada" & data$year==election),c("events")] <- 9
data[which(data$state=="New Hampshire" & data$year==election),c("events")] <- 9
data[which(data$state=="New Mexico" & data$year==election),c("events")] <- 9
data[which(data$state=="New York" & data$year==election),c("events")] <- 10
data[which(data$state=="North Carolina" & data$year==election),c("events")] <- 13
data[which(data$state=="Ohio" & data$year==election),c("events")] <- 35
data[which(data$state=="Pennsylvania" & data$year==election),c("events")] <- 29
data[which(data$state=="Tennessee" & data$year==election),c("events")] <- 1
data[which(data$state=="Texas" & data$year==election),c("events")] <- 1
data[which(data$state=="Virginia" & data$year==election),c("events")] <- 21
data[which(data$state=="West Virginia" & data$year==election),c("events")] <- 1
data[which(data$state=="Wisconsin" & data$year==election),c("events")] <- 1
```

From Hill, Rodriguez, and Wooden's "Stump Speeches and Road Trips: The Impact of State Campaign Appearances in Presidential Elections" we get the following data for 2004 and 2000: In 2004, AZ 11, AR 5, CA 10, CO 17, FL 97, GA 3, IA 55, ME 6, MI 37, MN 31, MO 28, NJ 5, NV 17, NH 15, NM 23, OH 112, OR 17, PA 61, SC 3, TN 5, VA 3, WA 5, WV 17, WI 73


```{r}
election <- 2004
data[which(data$state=="Arizona" & data$year==election),c("events")] <- 11
data[which(data$state=="Arkansas" & data$year==election),c("events")] <- 11
data[which(data$state=="California" & data$year==election),c("events")] <- 10
data[which(data$state=="Colorado" & data$year==election),c("events")] <- 17
data[which(data$state=="Florida" & data$year==election),c("events")] <- 97
data[which(data$state=="Georgia" & data$year==election),c("events")] <- 3
data[which(data$state=="Iowa" & data$year==election),c("events")] <- 55
data[which(data$state=="Maine" & data$year==election),c("events")] <- 6
data[which(data$state=="Michigan" & data$year==election),c("events")] <- 37
data[which(data$state=="Minnesota" & data$year==election),c("events")] <- 31
data[which(data$state=="Missouri" & data$year==election),c("events")] <- 28
data[which(data$state=="Nevada" & data$year==election),c("events")] <- 17
data[which(data$state=="New Hampshire" & data$year==election),c("events")] <- 15
data[which(data$state=="New Jersey" & data$year==election),c("events")] <- 5
data[which(data$state=="New Mexico" & data$year==election),c("events")] <- 23
data[which(data$state=="Ohio" & data$year==election),c("events")] <- 112
data[which(data$state=="Oregon" & data$year==election),c("events")] <- 17
data[which(data$state=="Pennsylvania" & data$year==election),c("events")] <- 61
data[which(data$state=="South Carolina" & data$year==election),c("events")] <- 3
data[which(data$state=="Tennessee" & data$year==election),c("events")] <- 5
data[which(data$state=="Virginia" & data$year==election),c("events")] <- 3
data[which(data$state=="Washington" & data$year==election),c("events")] <- 5
data[which(data$state=="West Virginia" & data$year==election),c("events")] <- 17
data[which(data$state=="Wisconsin" & data$year==election),c("events")] <- 73
```

In 2000 AZ 1, AR 15, CA 54, CO 6, FL 82, GA 6, IA 35, IL 35, KY 9, LA 12, ME 9, MI 56, MN 9, MO 29, NV 7, NH 9, NJ 10, NM 17, NY 25, NC 6, OH 35, OR 17, PA 56, TN 29, WA 18, WV 4, WI 37


```{r}
election <- 2000
data[which(data$state=="Arizona" & data$year==election),c("events")] <- 1
data[which(data$state=="Arkansas" & data$year==election),c("events")] <- 15
data[which(data$state=="California" & data$year==election),c("events")] <- 54
data[which(data$state=="Colorado" & data$year==election),c("events")] <- 6
data[which(data$state=="Florida" & data$year==election),c("events")] <- 82
data[which(data$state=="Georgia" & data$year==election),c("events")] <- 6
data[which(data$state=="Iowa" & data$year==election),c("events")] <- 35
data[which(data$state=="Kentucky" & data$year==election),c("events")] <- 9
data[which(data$state=="Louisiana" & data$year==election),c("events")] <- 12
data[which(data$state=="Maine" & data$year==election),c("events")] <- 9
data[which(data$state=="Michigan" & data$year==election),c("events")] <- 56
data[which(data$state=="Minnesota" & data$year==election),c("events")] <- 9
data[which(data$state=="Missouri" & data$year==election),c("events")] <- 29
data[which(data$state=="Nevada" & data$year==election),c("events")] <- 7
data[which(data$state=="New Hampshire" & data$year==election),c("events")] <- 9
data[which(data$state=="New Jersey" & data$year==election),c("events")] <- 10
data[which(data$state=="New Mexico" & data$year==election),c("events")] <- 17
data[which(data$state=="New York" & data$year==election),c("events")] <- 25
data[which(data$state=="North Carolina" & data$year==election),c("events")] <- 6
data[which(data$state=="Ohio" & data$year==election),c("events")] <- 35
data[which(data$state=="Oregon" & data$year==election),c("events")] <- 17
data[which(data$state=="Pennsylvania" & data$year==election),c("events")] <- 56
data[which(data$state=="Tennessee" & data$year==election),c("events")] <- 29
data[which(data$state=="Washington" & data$year==election),c("events")] <- 18
data[which(data$state=="West Virginia" & data$year==election),c("events")] <- 4
data[which(data$state=="Wisconsin" & data$year==election),c("events")] <- 37
```

## Logisic Regression

```{r}
new_train_data <- data[which(data$state %in% c("Colorado", 'Florida', 'Iowa', 'Michigan', "Nevada", 'New Hampshire', "New Mexico", 'Ohio', 'Oregon', 'Pennsylvania', "Wisconsin") & data$year == 2008),] %>% dplyr::select(-events)
new_train_data$is_swing <- 1
new_train_data <- inner_join(new_train_data, 
                             data[which((data$state %in% new_train_data$state) & (data$year == 2004)),c("state", "events", "margin")] %>% rename(previous_margin = margin),
                             by="state")
train_data <- new_train_data;

new_train_data <- data[which(data$state %in% c("Arizona", 'Arkansas', "California", "Georgia", "Indiana", "Kentucky", "Maine", 'Missouri', "New York", "North Carolina", "Tennessee", "Washington", "West Virginia", "Virginia") & data$year == 2008),] %>% dplyr::select(-events)
new_train_data$is_swing <- 0
new_train_data <- inner_join(new_train_data, 
                             data[which((data$state %in% new_train_data$state) & (data$year == 2004)),c("state", "events", "margin")] %>% rename(previous_margin = margin),
                             by="state")

train_data <- rbind(train_data,new_train_data);



new_train_data <- data[which(data$state %in% c("Colorado", 'Florida', 'Iowa', 'Michigan', 'Missouri', "Nevada", 'New Hampshire', "New Mexico", 'Ohio', 'Oregon', 'Pennsylvania', "West Virginia", "Wisconsin") & data$year == 2004),] %>% dplyr::select(-events)
new_train_data$is_swing <- 1
new_train_data <- inner_join(new_train_data,
                             data[which((data$state %in% new_train_data$state) & (data$year == 2000)),c("state", "events", "margin")] %>% rename(previous_margin = margin),
                             by="state")
train_data <- rbind(train_data,new_train_data);

new_train_data <- data[which(data$state %in% c("Arizona", "Arkansas", "California", "Georgia", "Maine", "New Jersey", "New Mexico", "South Carolina", "Tennessee", "Virginia", "Washington") & data$year == 2004),] %>% dplyr::select(-events)
new_train_data$is_swing <- 0

new_train_data <- inner_join(new_train_data,
                             data[which((data$state %in% new_train_data$state) & (data$year == 2000)),c("state", "events", "margin")] %>% rename(previous_margin = margin),
                             by="state")
train_data <- rbind(train_data,new_train_data);

train_data$mu <- 1/train_data$margin
```

We can then train a Bayesian logistic function

```{r}
# logit_mod <- step(bayesglm(is_swing ~ swings * margin * bellwether * events + previous_margin, data = train_data, family = binomial(link = "logit")));
logit_mod <- bayesglm(is_swing ~ swings + bellwether + events + previous_margin, data = train_data, family = binomial(link = "logit"));
lm0 <- step(glm(is_swing ~ swings + bellwether + events + previous_margin, data = train_data, family = binomial(link = "logit")));
weighted_logit_mod <- bayesglm(is_swing ~ weighted_swings + bellwether + events + previous_margin, data = train_data, family = binomial(link = "logit"));
weighted_lm0 <- step(glm(is_swing ~ weighted_swings + bellwether + events + previous_margin, data = train_data, family = binomial(link = "logit")));
```

## Retrodict 2012

Using just the number of swings, the margin of victory, and if it's a bellwether state, we have the following predictions:

```{r}
df_2012 <- inner_join(filter(data, year==2012) %>% dplyr::select(-events),
                      data %>%
                        filter(year == 2008) %>%
                        dplyr::select(margin, state, events) %>%
                        rename(previous_margin = margin),
                      by="state")

df_2012$bayes_predict <- predict(logit_mod, df_2012, type="response")
df_2012$predict <- predict(lm0, df_2012, type="response")

df_2012$weighted_bayes_predict <- predict(weighted_logit_mod, df_2012, type="response")
df_2012$weighted_predict <- predict(weighted_lm0, df_2012, type="response")

kable(df_2012 %>% dplyr::select(state,predict, bayes_predict, weighted_predict, weighted_bayes_predict) %>% arrange(-predict), row.names = FALSE)
```


## Retrodict 2016

Using just the number of swings, the margin of victory, and if it's a bellwether state, we have the following predictions:

```{r}
df_2016 <- inner_join(filter(data, year==2016) %>% dplyr::select(-events),
                      data %>%
                        filter(year == 2012) %>%
                        dplyr::select(margin, state, events) %>%
                        rename(previous_margin = margin),
                      by="state")

df_2016$bayes_predict <- predict(logit_mod, df_2016, type="response")
df_2016$predict <- predict(lm0, df_2016, type="response")
df_2016$weighted_bayes_predict <- predict(weighted_logit_mod, df_2016, type="response")
df_2016$weighted_predict <- predict(weighted_lm0, df_2016, type="response")
kable(df_2016 %>% dplyr::select(state,predict, bayes_predict, weighted_predict, weighted_bayes_predict) %>% arrange(-predict), row.names = FALSE)
```
## Predicting 2020

```{r}
df_2020 <- filter(data, year==2016) %>%
  rename(previous_margin = margin)

df_2020$bayes_predict <- predict(logit_mod, df_2020, type="response")
df_2020$predict <- predict(lm0, df_2020, type="response")

df_2020$weighted_bayes_predict <- predict(weighted_logit_mod, df_2020, type="response")
df_2020$weighted_predict <- predict(weighted_lm0, df_2020, type="response")

kable(df_2020 %>% dplyr::select(state, predict, bayes_predict, weighted_predict, weighted_bayes_predict) %>% arrange(-predict))
```

# Data Exploration

A histogram plot of the variances for the party vote's in each state.

```{r, fig.width=6}
path <- "../data/elections/presidential/state/1976_2016_president.RData"
results <- load_obj(path);
results <- make_party_into_factor(results);
results <- add_vote_percentage(results);
party_var <- results %>% group_by(state,party) %>% summarize(votepercentage_var = sd(votepercentage, na.rm=TRUE))
d1 <- party_var[which(party_var$party=='democrat'),c("votepercentage_var")]
hist(d1[["votepercentage_var"]],xlab="Percent of State's Vote",main="Democratic Vote Result St.Dev. by state, 1976-2016")
```


Similarly for the Republicans:


```{r, fig.width=6}
r1 <- party_var[which(party_var$party=='republican'),c("votepercentage_var")]
hist(r1[["votepercentage_var"]],xlab="Percent of State's Vote",main="Republican Vote Result St.Dev. by state, 1976-2016")
```




This is a fascinating property, Democratic voters seem to be solidly Democratic?