---
title: "Swingiest Counties in 2016"
author: "Alex Nelson"
date: "5/13/2019"
output:
  md_document:
    variant: gfm
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(dplyr)
library(car) # for vif()
library(usmap)
library(urbnmapr)
library(tidyverse)
library(ggplot2)
library(scales)
library(rmarkdown)
knitr::opts_chunk$set(echo = TRUE)
```

# Loading our data

We need a few helper functions to denormalize useful data in the rows.

```{r}
winner_by_state_and_year <- function(results) {
  return(results %>% group_by(state,year) %>% filter(votepercentage==max(votepercentage)) %>%  arrange(party));
}
```

We need to normalize the parties, since the Democratic party in Minnessota calls themselves the Democratic-Farmer-Labor party.

```{r}
make_party_into_factor <- function(results) {
  results$party[results$party == 'democratic-farmer-labor'] <- 'democrat';
  results$party[is.na(results$party)] <- '$third-party';
  results$party <- as.factor(results$party);
  return(results);
}
add_vote_percentage <- function(results) {
  results$votepercentage <- with(results, candidatevotes/totalvotes);
  return(results);
}
```

We need a helper function to load `RData` files into specified variables.

```{r}
load_obj <- function(path) {
  env <- new.env();
  nm <- load(path,env)[1];
  env[[nm]];
}
```

```{r}
count_alternates <- function(column) {
  count <- 0;
  if (length(column) > 0) {
    val <- column[[1]][1];
    for (i in 2:length(column)) {
      if (column[[i]][1] != val) {
        val <- column[[i]][1];
        count <- count + 1;
      }
    }
  }
  return(count);
}
```

```{r}
number_of_swings_since <- function(results, election_year, state_name, number_of_elections=7) {
  start_year <- election_year - (number_of_elections - 1)*4;
  has_data <- (1976 <= start_year);
  is_election_year <- ((election_year - 1976)%%4 == 0);
  if (!has_data || !is_election_year) {
    return(0);
  } else {
    parties<-results[which(results$state==state_name & start_year <= results$year & results$year <= election_year),c("party")];
    return(count_alternates(parties[["party"]]));
  }
}

has_swung_since <- function(results, election_year, state_name, number_of_elections=7) {
  return(number_of_swings_since(results, election_year, state_name, number_of_elections) > 1);
}
```

```{r}
state_path <- "../data/elections/presidential/state/1976_2016_president.RData"
state_results <- load_obj(state_path);
state_results <- make_party_into_factor(state_results);
state_results <- add_vote_percentage(state_results);
state_winner_data <- winner_by_state_and_year(state_results);
state_winner_data <- state_winner_data[with(state_winner_data,order(state,year)), ]

swing_data <- state_winner_data[which(state_winner_data$year >= 2000),c("state","year")]
```

```{r}
for (i in 1:nrow(swing_data)) {
  row <- swing_data[i,];
  year <- row[['year']];
  state <- row[['state']];
  swing_data[which(swing_data$state == state & swing_data$year == year),c('swings')] <- number_of_swings_since(state_winner_data,year,state,2);
}
```

The states we are interested in examining are the ones which have swung from Obama to Trump.

```{r}
kable(swing_data[which(swing_data$year==2016 & swing_data$swings>=1),])
```

```{r}
path <- "../data/elections/presidential/county/countypres_2000-2016.RData"
results <- load_obj(path);
results <- make_party_into_factor(results);
results <- add_vote_percentage(results);
winner_data <- winner_by_state_and_year(results);
winner_data <- winner_data[with(winner_data,order(state,county,year)), ]
```

We now want to drill down on the states which were won by Obama but switched to Trump.

```{r}
states <- swing_data[which(swing_data$year==2016 & swing_data$swings>=1),c('state')]$state
relevant_data <- results[results$state %in% states,]
```

We want to find, among these states, which counties were won by Obama but switched to Trump. We don't know if comparing 2008 to 2016 is more meaningful than 2012 to 2016, so we gather all the data we can.

```{r}
counties_obama_won_2008 <- relevant_data %>% filter(year==2008) %>% group_by(state,county) %>% filter(votepercentage==max(votepercentage)) %>%  arrange(party) %>% filter(party=='democrat')
counties_obama_won_2008 <- add_vote_percentage(counties_obama_won_2008)

counties_obama_won_2012 <- relevant_data %>% filter(year==2012) %>% group_by(state,county) %>% filter(votepercentage==max(votepercentage)) %>%  arrange(party) %>% filter(party=='democrat')
counties_obama_won_2012 <- add_vote_percentage(counties_obama_won_2012)

counties_trump_won <- relevant_data %>% filter(year==2016) %>% group_by(state,county) %>% filter(votepercentage==max(votepercentage)) %>%  arrange(party) %>% filter(party=='republican')
counties_trump_won <- add_vote_percentage(counties_trump_won)
```

Now we are curious about the counties which swung from Obama to Trump, which we will try measuring in terms of percent of votes cast.

```{r}
trump_intersect_2008 <- counties_trump_won[counties_trump_won$state %in% counties_obama_won_2008$state & counties_trump_won$county %in% counties_obama_won_2008$county,]
obama_intersect_2008 <- counties_obama_won_2008[counties_obama_won_2008$state %in% counties_trump_won$state & counties_obama_won_2008$county %in% counties_trump_won$county,]

compare_2008 <- merge(obama_intersect_2008, trump_intersect_2008, by=c('state','county','FIPS'))
compare_2008$trump_gain <- compare_2008$votepercentage.y - compare_2008$votepercentage.x

data_2008 <- compare_2008[which(compare_2008$trump_gain > 0),c('state','county','FIPS','trump_gain')]
```

```{r}
kable(data_2008[order(-data_2008$trump_gain),])
```
# Plots for the 2008 Results

Now we can try to plot the data:

```{r}
data_2008 <- compare_2008[,c('state','county','FIPS','trump_gain')]
county_2008 <- data_2008
county_2008$county_fips = as.character(county_2008$FIPS)
county_2008 <- left_join(counties, county_2008, by = "county_fips") 

county_2008 %>% 
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = NA) +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = 'white',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

We plot each state independently

## Florida

```{r}
county_2008 %>% filter(state_name == 'Florida') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```


```{r}
fl_2008 <- (data_2008 %>% filter(state == 'Florida'))
kable(fl_2008[order(-fl_2008$trump_gain),c('county','trump_gain')])
```

## Iowa

```{r}
county_2008 %>% filter(state_name == 'Iowa') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
ia_2008 <- (data_2008 %>% filter(state == 'Iowa'))
kable(ia_2008[order(-ia_2008$trump_gain),c('county','trump_gain')])
```

## Michigan


```{r}
county_2008 %>% filter(state_name == 'Michigan') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
mi_2008 <- (data_2008 %>% filter(state == 'Michigan'))
kable(mi_2008[order(-mi_2008$trump_gain),c('county','trump_gain')])
```

## Nebraska

```{r}
county_2008 %>% filter(state_name == 'Nebraska') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
ne_2008 <- (data_2008 %>% filter(state == 'Nebraska'))
kable(ne_2008[order(-ne_2008$trump_gain),c('county','trump_gain')])
```

## Ohio

```{r}
county_2008 %>% filter(state_name == 'Ohio') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
oh_2008 <- (data_2008 %>% filter(state == 'Ohio'))
kable(oh_2008[order(-oh_2008$trump_gain),c('county','trump_gain')])
```

## Pennsylvania

```{r}
county_2008 %>% filter(state_name == 'Pennsylvania') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
pa_2008 <- (data_2008 %>% filter(state == 'Pennsylvania'))
kable(pa_2008[order(-pa_2008$trump_gain),c('county','trump_gain')])
```

## Wisconsin


```{r}
county_2008 %>% filter(state_name == 'Wisconsin') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
wi_2008 <- (data_2008 %>% filter(state == 'Wisconsin'))
kable(wi_2008[order(-wi_2008$trump_gain),c('county','trump_gain')])
```

# Plots for the 2012 Results

And the comparison for 2012 results


```{r}
trump_intersect_2012 <- counties_trump_won[counties_trump_won$state %in% counties_obama_won_2012$state & counties_trump_won$county %in% counties_obama_won_2012$county,]
obama_intersect_2012 <- counties_obama_won_2012[counties_obama_won_2012$state %in% counties_trump_won$state & counties_obama_won_2012$county %in% counties_trump_won$county,]

compare_2012 <- merge(obama_intersect_2012, trump_intersect_2012, by=c('state','county','FIPS'))
compare_2012$trump_gain <- compare_2012$votepercentage.y - compare_2012$votepercentage.x

data_2012 <- compare_2012[which(compare_2012$trump_gain > 0),c('state','county','FIPS','trump_gain')]
```

There were `nrow(data_2012)` counties which flipped from Obama 2012 to Trump in 2016:

```{r}
kable(data_2012[order(-data_2012$trump_gain),])
```

Now we can try to plot the data:

```{r}
data_2012 <- compare_2012[,c('state','county','FIPS','trump_gain')]
county_2012 <- data_2012
county_2012$county_fips = as.character(county_2012$FIPS)
county_2012 <- left_join(counties, county_2012, by = "county_fips")

county_2012 %>% 
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = NA) +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

## Florida

```{r}
county_2012 %>% filter(state_name == 'Florida') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```


```{r}
fl_2012 <- (data_2012 %>% filter(state == 'Florida'))
kable(fl_2012[order(-fl_2012$trump_gain),c('county','trump_gain')])
```


## Iowa

```{r}
county_2012 %>% filter(state_name == 'Iowa') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
ia_2012 <- (data_2012 %>% filter(state == 'Iowa'))
kable(ia_2012[order(-ia_2012$trump_gain),c('county','trump_gain')])
```

## Michigan

```{r}
county_2012 %>% filter(state_name == 'Michigan') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
mi_2012 <- (data_2012 %>% filter(state == 'Michigan'))
kable(mi_2012[order(-mi_2012$trump_gain),c('county','trump_gain')])
```

## Ohio

```{r}
county_2012 %>% filter(state_name == 'Ohio') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
oh_2012 <- (data_2012 %>% filter(state == 'Ohio'))
kable(oh_2012[order(-oh_2012$trump_gain),c('county','trump_gain')])
```

## Pennsylvania

```{r}
county_2012 %>% filter(state_name == 'Pennsylvania') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
pa_2012 <- (data_2012 %>% filter(state == 'Pennsylvania'))
kable(pa_2012[order(-pa_2012$trump_gain),c('county','trump_gain')])
```

## Wisconsin

```{r}
county_2012 %>% filter(state_name == 'Wisconsin') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
wi_2012 <- (data_2012 %>% filter(state == 'Wisconsin'))
kable(wi_2012[order(-wi_2012$trump_gain),c('county','trump_gain')])
```




