---
title: "Central Limit Theorem"
author: "Alex Nelson"
date: "6/18/2019"
output: 
  md_document:
    variant: gfm
    toc: true
    toc_depth: 6
---

```{r setup, include=FALSE}
library(rmarkdown)
library(kableExtra)

library(ggplot2)
library(ggthemes)

knitr::opts_chunk$set(echo = TRUE)
```

This is a quick attempt to reproduce the results from:

+ Zachary R. Smith and Craig S. Wells's [Central Limit Theorem and Sample Size](https://www.umass.edu/remp/Papers/Smith&Wells_NERA06.pdf)

We begin with three probability distributions:  Normal (`dnorm(mean = 50, sd = 10)`)), Uniform
(`dunif(10,30)`), Bimodal, and then we will consider a family of skewed distributions. 

> From each of the three populations, 10,000 samples of size 5, 10, 15, 20, 25, and 30 were randomly drawn, rounding to the nearest integer, as is often observed in educational and psychological measures.  The mean for each sample was then computed, using four decimal places. Twenty replications were performed for each condition (i.e., the process was repeated 20 times for each sample size).
> 
> After the sampling distribution for the mean was constructed for a replication, a one sample Kolmogorov-Smirnov (KS) test was used to test whether the sample mean followed a normal distribution.  This test was used on each sample size taken from all distributions.  The proportion of replications in which the distribution of the sample means were concluded to depart normality according to the KS-test was recorded. 

So lets start sampling in the manner described.

```{r}
sample_normal <- function(sample_size, N = 20, mean = 50, sd = 10) {
  replicate(N, mean(round(rnorm(sample_size, mean, sd))));
}

sample_unif <- function(sample_size, N=20, lower_bound = 10, upper_bound = 30) {
  replicate(N, mean(round(runif(sample_size, lower_bound, upper_bound))))
}

sample_bimodal <- function(sample_size, N=20, theta = 0.75) {
  samples = c();
  for (i in 1:N) {
    w = rbinom(sample_size, 1, theta);
    samples <- append(samples, mean(round(w*rnorm(sample_size, 50, 10) + (1 - w)*rnorm(sample_size, 85, 10))));
  }
  samples
}

fleisman <- function(n, a, b, c, d, mean = 50, sd = 10) {
  x <- rnorm(n, mean, sd)
  a + x*(b + x*(c + d*x))
}
coefs <- list(list(-0.399, 0.930, 0.399, -0.036),
              list(-0.221, 0.866, 0.221, 0.027),
              list(-0.161, 0.819, 0.161, 0.049),
              list(-0.119, 0.789, 0.119, 0.062));


sample_fleisman <- function(sample_size, k, N=20, theta = 0.75) {
  a <- coefs[[k]][[1]]
  b <- coefs[[k]][[2]]
  c <- coefs[[k]][[3]]
  d <- coefs[[k]][[4]]
  replicate(N, mean(round(fleisman(sample_size, a, b, c, d))));
}
```

# Plotting 

Before continuing on, we can plot some histograms to see what the data looks like.

```{r}
qplot(sample_bimodal(5, N=200), geom="histogram", fill=I("#008FD5"), ylab="Bimodal(5, N=200)") +
  scale_color_fivethirtyeight() +
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) + xlab("Bimodal(5, N=200)") + ylab("density")
```

```{r}
qplot(sample_bimodal(10, N=200), geom="histogram", fill=I("#008FD5")) +
  scale_color_fivethirtyeight() +
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) + xlab("Bimodal(10, N=200)") + ylab("density")
```

```{r}
qplot(sample_bimodal(30, N=200), geom="histogram", fill=I("#008FD5")) +
  scale_color_fivethirtyeight() +
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) + xlab("Bimodal(30, N=200)") + ylab("density")
```

```{r}
qplot(sample_fleisman(5, 4, N=200), geom="histogram", fill=I("#FF2700")) +
  scale_color_fivethirtyeight() +
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) + xlab("Fleisman4(5, N=200)") + ylab("density")
```

# Testing

Perfect, now we are in a state to imitate the testing procedure:

> Samples of 5 through 300 at intervals of 5 were drawn with no rounding from each of the four skewed distributions over 10,000 repetitions. After the transformation, the sampling means were rounded to the fourth decimal place. Once the sampling distribution for the mean is complete, the one sample Kolmogorov-Smirnov (KS) test was applied.


We implement this thus:

```{r}
results <- data.frame(uniform = rep(NA,6),
                      normal = rep(NA,6),
                      bimodal = rep(NA,6),
                      fleisman1 = rep(NA,6),
                      fleisman2 = rep(NA,6),
                      fleisman3 = rep(NA,6),
                      fleisman4 = rep(NA,6))
# sidak significance = 1 - (1 - 0.05)**(1e-4);
significance_level <- 0.05;
for (i in 1:6) {
  print(i)
  results$uniform[i] <- length(Filter(function(p) {p < significance_level}, replicate(10000, shapiro.test(sample_unif(5*i))$p)))
  results$normal[i] <- length(Filter(function(p) {p < significance_level}, replicate(10000, shapiro.test(sample_normal(5*i))$p)))
  results$bimodal[i] <- length(Filter(function(p) {p < significance_level}, replicate(10000, shapiro.test(sample_bimodal(5*i))$p)))
  results$fleisman1[i] <- length(Filter(function(p) {p < significance_level}, replicate(10000, shapiro.test(sample_fleisman(5*i, 1))$p)))
  results$fleisman2[i] <- length(Filter(function(p) {p < significance_level}, replicate(10000, shapiro.test(sample_fleisman(5*i, 2))$p)))
  results$fleisman3[i] <- length(Filter(function(p) {p < significance_level}, replicate(10000, shapiro.test(sample_fleisman(5*i, 3))$p)))
  results$fleisman4[i] <- length(Filter(function(p) {p < significance_level}, replicate(10000, shapiro.test(sample_fleisman(5*i, 4))$p)))
}
```

```{r}
kable(results)
```

This is nothing at all what Smith and Wells report...

# Conclusion

The central limit theorem works, the world is safe, and we can all sleep well again. AND DO NOT CITE THIS DAMN PAPER!