---
title: "Democratic Decision Tree"
author: "Alex Nelson"
date: "6/19/2019"
output: 
  md_document:
    variant: gfm
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)                # data cleaning and ggplot2 - tidy!
library(stringr)                  # string manipulation
library(tibble)
library(party)
library(rpart)
library(rpart.plot)
library(rmarkdown)
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

I want to create a decision tree to guide the user to determine who is likely to win the nomination. Fortunately, I have created a tidy dataset for the Democratic primaries, so lets load it up.

```{r, load_tidy_data}
primary_data <- rbind(read_csv("../data/primary/democrat_1984.csv"),
                      read_csv("../data/primary/democrat_1988.csv"),
                      read_csv("../data/primary/democrat_1992.csv"),
                      read_csv("../data/primary/democrat_2004.csv"),
                      read_csv("../data/primary/democrat_2008.csv")) %>%
  group_by(state,event_type) %>%
  filter(any(won_state),!is.na(date)) %>%
  ungroup() %>%
  filter(!is.na(date)) %>%
  mutate(year = format(date,"%Y"),
         won_nomination = ifelse(is_nominee, 1, 0),
         event_type = factor(event_type, labels = c("primary","caucus"))) %>%
  group_by(year) %>%
  mutate(days_since_start = as.numeric(date - min(date))) %>%
  ungroup() %>%
  select(-year)
```


```{r}
dt1 <- primary_data
dt1$state <- tolower(str_replace_all(dt1$state, "\\s+", "_"))
dt1$state <- str_replace(dt1$state, "\\[\\d+\\]", "")
dt1$year <- format(dt1$date, "%Y")


dt2 <- dt1 %>% 
  mutate(won_state = ifelse(won_state, 1, 0)) %>%
  filter(days_since_start < 20) %>%
  select(candidate,year,state,won_nomination,won_state) %>%
  rowid_to_column() %>%
  group_by(candidate,year) %>%
  nest() %>%
  mutate(data = map(data, spread, state, won_state, fill = 0)) %>%
  unnest() %>% select(-rowid) %>% group_by(candidate,year,won_nomination) %>% summarise_each(funs(max))
```

We can now construct the decision tree. The result is rather interesting. It is folklore that New Hampshire is only relevant when trying to weed out a sitting President in a primary challenge (like Teddy Kennedy challenging Carter in 1980), and all other circumstances is irrelevant. We find this reflected in the decision tree:


```{r}
rtree_fit <- rpart(won_nomination ~ iowa + minnesota + new_hampshire + south_dakota + maine + district_of_columbia, 
                   dt2, method = "class", model = T, 
                   parms = list(prior = c(.65,.35), split = "information"),
                   control = rpart.control(minsplit = 2)
                   )
rpart.plot(rtree_fit)
```

We should read this tree as stating three rows in each outcome bubble:

1. Did the candidate become the nominee? `0` for "No", `1` for "Yes"
2. The probability of this outcome (0.00 for "impossible to become nominee" to 1.00 for "almost certainly the nominee")
3. The percentage of data points which fall in the outcome.

In other words, the unsurprising outcome discerned: to become the nominee, the candidate must win _either_ Iowa _or_ New Hampshire.

## Logistic Regression

We filter data, in case we accidentally allowed bad data through.

```{r}
dt0 <- filter(primary_data, !is.na(vote_perc))
```

Now we construct a few logistic regressions.

```{r}
logitMod <- glm(won_nomination ~ won_state*days_since_start*vote_perc, 
                data = dt0, family = binomial(link = "logit"))
summary(logitMod)
```

```{r}
summary(glm(won_nomination ~ won_state:vote_perc:days_since_start,
            data = dt0, family = binomial(link = "logit")))
```


```{r}
predicted <- predict(logitMod, dt0, type = "response")  # predicted scores
dt0$prediction <- predicted

precision <- (dt0 %>% filter(is_nominee) %>% select(prediction) %>% sum)/sum(predicted)
recall <- (dt0 %>% filter(is_nominee) %>% select(prediction) %>% sum)/(dt0 %>% filter(is_nominee) %>% nrow)
Fmeasure <- 2 * precision * recall / (precision + recall)
print(Fmeasure)
```


