---
title: "Democratic Decision Tree"
author: "Alex Nelson"
date: "6/19/2019"
output: 
  md_document:
    variant: gfm
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)                # data cleaning and ggplot2 - tidy!
library(stringr)                  # string manipulation
library(tibble)
library(party)
library(rpart)
library(rpart.plot)
library(rmarkdown)
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

I want to create a decision tree to guide the user to determine who is likely to win the nomination.

```{r}
.data <- as.data.frame(tribble(
  ~name,    ~year, ~iowa, ~new_hampshire, ~nevada, ~south_carolina, ~won_nomination,
  "Clinton", 2016,     1,              0,       1,               1,               1,
  "Sanders", 2016,     0,              1,       0,               0,               0,
  "Obama",   2008,     1,              0,       1,               1,               1,
  "Edwards", 2008,     0,              0,       0,               0,               0,
  "Clinton", 2008,     0,              0,       0,               0,               0,
  "Kerry",   2004,     1,              1,       1,               0,               1,
  "Edwards", 2004,     1,              0,       0,               1,               0,
  "Harkin",  1992,     1,              0,       0,               0,               0,
  "Tsongas", 1992,     0,              1,       1,               0,               0,
  "Clinton", 1992,     0,              1,       1,               1,               1
))
```

```{r}
tree_fit <- ctree(won_nomination ~ iowa + new_hampshire + nevada + south_carolina, data = .data)
```

```{r}
rtree_fit <- rpart(won_nomination ~ iowa + new_hampshire + nevada + south_carolina, .data, method = "class", model = T, control = rpart.control(minsplit = 2))
rpart.plot(rtree_fit)
```

We should read this tree as stating three rows in each outcome bubble:

1. Did the candidate become the nominee? `0` for "No", `1` for "Yes"
2. The probability of this outcome (0.00 for "impossible to become nominee" to 1.00 for "almost certainly the nominee")
3. The percentage of data points which fall in the outcome.