---
title: "How is my candidate doing?"
author: "Alex Nelson"
date: "6/22/2019"
output:
  md_document:
    variant: gfm
    toc: true
    toc_depth: 6
---

```{r setup, include=FALSE}
library(kableExtra)

library(tidyverse)

library(assertthat)

library(rmarkdown)
knitr::opts_chunk$set(echo = TRUE)
```

The first thing we do is load the 538 collected poll data.

```{r}
poll_df <- read_csv("../data/primary/2020/president_primary_polls.csv")
poll_df$start_date <- as.Date(poll_df$start_date, "%m/%d/%y")
poll_df$end_date <- as.Date(poll_df$end_date, "%m/%d/%y")
poll_2020_df <- filter(poll_df, cycle==2020, party=="DEM")
```

Now we can arrange by the `end_date` for the poll, which places the earlier polls first and the more recent polls at the end of the data frame. This sets us up nicely to use a running average for the poll for each candidate.

# Vanilla moving average

We implement the "naive" moving average.

```{r, moving average}
alpha <- function(prev_date, next_date, avg_interval, normalizer = 30, ...) {
  1 - exp(-abs(as.numeric(next_date - prev_date))/(normalizer * avg_interval))
}
```

We will have a utility function for computing the average interval between poll releases.

```{r}
average_interval <- function(poll_data) {
  abs(mean(as.numeric(diff(poll_data$end_date))))
}
```

The moving average is a little sad, it's just what we'd expect.

```{r}
moving_average <- function(poll_data, weighting_interval, alpha_fn = alpha) {
  assert_that(length(unique(poll_data$cycle)) == 1)
  poll_data < arrange(poll_data, end_date)
  avg_interval <- average_interval(poll_data)
  avg <- poll_data$pct[1]
  prev_dt <- poll_data$end_date[1]
  for (i in 2:nrow(poll_data)) {
    a <- alpha_fn(prev_dt, poll_data$end_date[i], avg_interval, weighting_interval, poll_data$fte_grade[i])
    avg <- a*poll_data$pct[i] + (1 - a)*avg
    prev_dt <- poll_data$end_date[i]
  }
  avg
}

compute_moving_avgs <- function(poll_data, weighting_interval = 30) {
  assert_that(length(unique(poll_data$cycle)) == 1)
  
  poll_data %>%
    group_by(candidate_name) %>%
    filter(n() > 2, !is.na(fte_grade), !is.na(sponsors), state=="Iowa") %>%
    nest() %>%
    transmute(candidate_name,
              moving_avg = map(data, moving_average, weighting_interval)
              ) %>%
    unnest() %>%
    arrange(-moving_avg)
}
```

We can now present a table of running averages by month:

```{r}
kable(compute_moving_avgs(poll_2020_df))
```
And if we choose instead a week:


```{r}
kable(compute_moving_avgs(poll_2020_df, 7))
```


```{r}
kable(compute_moving_avgs(poll_2020_df, 1))
```

```{r}
kable(compute_moving_avgs(poll_2020_df, 45))
```

# Weighing the score of the pollsters

There are a lot of garbage polls, so we need to apply a weighting scheme to them.

```{r}
grade_pollster <- function(fte_grade) {
  lookup <- list("A+" = 13/13,
                 "A"  = 12/13,
                 "A-" = 11/13,
                 "B+" = 10/13,
                 "B"  =  9/39,
                 "B-" =  8/39,
                 "C+" =  6/39,
                 "C"  =  5/39,
                 "C-" =  4/39,
                 "D+" =  3/39,
                 "D"  =  2/39,
                 "D-" =  1/39)
  ifelse(is.null(lookup[[fte_grade]]), 0, lookup[[fte_grade]])
}
```

Now we apply this to our new alpha function.

```{r}
weighted_alpha <- function(prev_date, next_date, avg_interval, normalizer, fte_grade, ...) {
  grade_pollster(fte_grade)*alpha(prev_date, next_date, avg_interval, normalizer)
}

compute_weighted_moving_avgs <- function(poll_data, weighting_interval = 30) {
  assert_that(length(unique(poll_data$cycle)) == 1)
  
  poll_data %>%
    group_by(candidate_name) %>%
    filter(n() > 2, !is.na(fte_grade), !is.na(sponsors), state=="Iowa") %>%
    nest() %>%
    transmute(candidate_name,
              moving_avg = map(data, moving_average, weighting_interval, weighted_alpha)
              ) %>%
    unnest() %>%
    arrange(-moving_avg)
}

kable(compute_weighted_moving_avgs(poll_2020_df, 7))
```

It seems the larger the window for remembering events, the smaller the lead Biden has, until we weigh the pollster ratings.

```{r}
kable(compute_weighted_moving_avgs(poll_2020_df))
```

