---
title: "Swingiest Counties in 2016"
author: "Alex Nelson"
date: "5/13/2019"
output:
  md_document:
    variant: gfm
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)

library(dplyr)
library(car) # for vif()
library(usmap)
library(urbnmapr)

library(arm) # for glmer

library(tidycensus)
library(tidyverse)
library(ggplot2)
library(scales) # for muted()
library(rmarkdown)
knitr::opts_chunk$set(echo = TRUE)
```

# Loading our data

We need a few helper functions to denormalize useful data in the rows.

```{r}
winner_by_state_and_year <- function(results) {
  return(results %>% group_by(state,year) %>% filter(votepercentage==max(votepercentage)) %>%  arrange(party));
}
```

We need to normalize the parties, since the Democratic party in Minnessota calls themselves the Democratic-Farmer-Labor party.

```{r}
make_party_into_factor <- function(results) {
  results$party[results$party == 'democratic-farmer-labor'] <- 'democrat';
  results$party[is.na(results$party)] <- '$third-party';
  results$party <- as.factor(results$party);
  return(results);
}
add_vote_percentage <- function(results) {
  results$votepercentage <- with(results, candidatevotes/totalvotes);
  return(results);
}
```

We need a helper function to load `RData` files into specified variables.

```{r}
load_obj <- function(path) {
  env <- new.env();
  nm <- load(path,env)[1];
  env[[nm]];
}
```

```{r}
count_alternates <- function(column) {
  count <- 0;
  if (length(column) > 0) {
    val <- column[[1]][1];
    for (i in 2:length(column)) {
      if (column[[i]][1] != val) {
        val <- column[[i]][1];
        count <- count + 1;
      }
    }
  }
  return(count);
}
```

```{r}
number_of_swings_since <- function(results, election_year, state_name, number_of_elections=7) {
  start_year <- election_year - (number_of_elections - 1)*4;
  has_data <- (1976 <= start_year);
  is_election_year <- ((election_year - 1976)%%4 == 0);
  if (!has_data || !is_election_year) {
    return(0);
  } else {
    parties<-results[which(results$state==state_name & start_year <= results$year & results$year <= election_year),c("party")];
    return(count_alternates(parties[["party"]]));
  }
}

has_swung_since <- function(results, election_year, state_name, number_of_elections=7) {
  return(number_of_swings_since(results, election_year, state_name, number_of_elections) > 1);
}
```

```{r}
state_path <- "../data/elections/presidential/state/1976_2016_president.RData"
state_results <- load_obj(state_path);
state_results <- make_party_into_factor(state_results);
state_results <- add_vote_percentage(state_results);
state_winner_data <- winner_by_state_and_year(state_results);
state_winner_data <- state_winner_data[with(state_winner_data,order(state,year)), ]

swing_data <- state_winner_data[which(state_winner_data$year >= 2000),c("state","year")]
```

```{r}
for (i in 1:nrow(swing_data)) {
  row <- swing_data[i,];
  year <- row[['year']];
  state <- row[['state']];
  swing_data[which(swing_data$state == state & swing_data$year == year),c('swings')] <- number_of_swings_since(state_winner_data,year,state,2);
}
```

The states we are interested in examining are the ones which have swung from Obama to Trump.

```{r}
kable(swing_data[which(swing_data$year==2016 & swing_data$swings>=1),])
```

```{r}
path <- "../data/elections/presidential/county/countypres_2000-2016.RData"
results <- load_obj(path);
results <- make_party_into_factor(results);
results <- add_vote_percentage(results);
county_results <- results; # for later usage
winner_data <- winner_by_state_and_year(results);
winner_data <- winner_data[with(winner_data,order(state,county,year)), ]
```

We now want to drill down on the states which were won by Obama but switched to Trump.

```{r}
states <- swing_data[which(swing_data$year==2016 & swing_data$swings>=1),c('state')]$state
relevant_data <- results[results$state %in% states,]
```

We want to find, among these states, which counties were won by Obama but switched to Trump. We don't know if comparing 2008 to 2016 is more meaningful than 2012 to 2016, so we gather all the data we can.

```{r}
counties_obama_won_2008 <- relevant_data %>% filter(year==2008) %>% group_by(state,county) %>% filter(votepercentage==max(votepercentage)) %>%  arrange(party) %>% filter(party=='democrat')
counties_obama_won_2008 <- add_vote_percentage(counties_obama_won_2008)

counties_obama_won_2012 <- relevant_data %>% filter(year==2012) %>% group_by(state,county) %>% filter(votepercentage==max(votepercentage)) %>%  arrange(party) %>% filter(party=='democrat')
counties_obama_won_2012 <- add_vote_percentage(counties_obama_won_2012)

counties_trump_won <- relevant_data %>% filter(year==2016) %>% group_by(state,county) %>% filter(votepercentage==max(votepercentage)) %>%  arrange(party) %>% filter(party=='republican')
counties_trump_won <- add_vote_percentage(counties_trump_won)
```

Now we are curious about the counties which swung from Obama to Trump, which we will try measuring in terms of percent of votes cast.

```{r}
trump_intersect_2008 <- counties_trump_won[counties_trump_won$state %in% counties_obama_won_2008$state & counties_trump_won$county %in% counties_obama_won_2008$county,]
obama_intersect_2008 <- counties_obama_won_2008[counties_obama_won_2008$state %in% counties_trump_won$state & counties_obama_won_2008$county %in% counties_trump_won$county,]

compare_2008 <- merge(obama_intersect_2008, trump_intersect_2008, by=c('state','county','FIPS'))
compare_2008$trump_gain <- compare_2008$votepercentage.y - compare_2008$votepercentage.x

data_2008 <- compare_2008[which(compare_2008$trump_gain > 0),c('state','county','FIPS','trump_gain')]
```

```{r}
kable(data_2008[order(-data_2008$trump_gain),])
```
# Plots for the 2008 Results

Now we can try to plot the data:

```{r}
data_2008 <- compare_2008[,c('state','county','FIPS','trump_gain')]
county_2008 <- data_2008
county_2008$county_fips = as.character(county_2008$FIPS)
county_2008 <- left_join(counties, county_2008, by = "county_fips") 

county_2008 %>% 
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = NA) +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = 'white',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

We plot each state independently

## Florida

```{r}
county_2008 %>% filter(state_name == 'Florida') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```


```{r}
fl_2008 <- (data_2008 %>% filter(state == 'Florida'))
kable(fl_2008[order(-fl_2008$trump_gain),c('county','trump_gain')])
```

## Iowa

```{r}
county_2008 %>% filter(state_name == 'Iowa') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
ia_2008 <- (data_2008 %>% filter(state == 'Iowa'))
kable(ia_2008[order(-ia_2008$trump_gain),c('county','trump_gain')])
```

## Michigan


```{r}
county_2008 %>% filter(state_name == 'Michigan') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
mi_2008 <- (data_2008 %>% filter(state == 'Michigan'))
kable(mi_2008[order(-mi_2008$trump_gain),c('county','trump_gain')])
```

## Nebraska

```{r}
county_2008 %>% filter(state_name == 'Nebraska') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
ne_2008 <- (data_2008 %>% filter(state == 'Nebraska'))
kable(ne_2008[order(-ne_2008$trump_gain),c('county','trump_gain')])
```

## Ohio

```{r}
county_2008 %>% filter(state_name == 'Ohio') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
oh_2008 <- (data_2008 %>% filter(state == 'Ohio'))
kable(oh_2008[order(-oh_2008$trump_gain),c('county','trump_gain')])
```

## Pennsylvania

```{r}
county_2008 %>% filter(state_name == 'Pennsylvania') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
pa_2008 <- (data_2008 %>% filter(state == 'Pennsylvania'))
kable(pa_2008[order(-pa_2008$trump_gain),c('county','trump_gain')])
```

## Wisconsin


```{r}
county_2008 %>% filter(state_name == 'Wisconsin') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
wi_2008 <- (data_2008 %>% filter(state == 'Wisconsin'))
kable(wi_2008[order(-wi_2008$trump_gain),c('county','trump_gain')])
```

# Plots for the 2012 Results

And the comparison for 2012 results


```{r}
trump_intersect_2012 <- counties_trump_won[counties_trump_won$state %in% counties_obama_won_2012$state & counties_trump_won$county %in% counties_obama_won_2012$county,]
obama_intersect_2012 <- counties_obama_won_2012[counties_obama_won_2012$state %in% counties_trump_won$state & counties_obama_won_2012$county %in% counties_trump_won$county,]

compare_2012 <- merge(obama_intersect_2012, trump_intersect_2012, by=c('state','county','FIPS'))
compare_2012$trump_gain <- compare_2012$votepercentage.y - compare_2012$votepercentage.x

data_2012 <- compare_2012[which(compare_2012$trump_gain > 0),c('state','county','FIPS','trump_gain')]
```

There were `nrow(data_2012)` counties which flipped from Obama 2012 to Trump in 2016:

```{r}
kable(data_2012[order(-data_2012$trump_gain),])
```

Now we can try to plot the data:

```{r}
data_2012 <- compare_2012[,c('state','county','FIPS','trump_gain')]
county_2012 <- data_2012
county_2012$county_fips = as.character(county_2012$FIPS)
county_2012 <- left_join(counties, county_2012, by = "county_fips")

county_2012 %>% 
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = NA) +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

## Florida

```{r}
county_2012 %>% filter(state_name == 'Florida') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```


```{r}
fl_2012 <- (data_2012 %>% filter(state == 'Florida'))
kable(fl_2012[order(-fl_2012$trump_gain),c('county','trump_gain')])
```


## Iowa

```{r}
county_2012 %>% filter(state_name == 'Iowa') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
ia_2012 <- (data_2012 %>% filter(state == 'Iowa'))
kable(ia_2012[order(-ia_2012$trump_gain),c('county','trump_gain')])
```

## Michigan

```{r}
county_2012 %>% filter(state_name == 'Michigan') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```
```{r}
mi_2012 <- (data_2012 %>% filter(state == 'Michigan'))
kable(mi_2012[order(-mi_2012$trump_gain),c('county','trump_gain')])
```

## Ohio

```{r}
county_2012 %>% filter(state_name == 'Ohio') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
oh_2012 <- (data_2012 %>% filter(state == 'Ohio'))
kable(oh_2012[order(-oh_2012$trump_gain),c('county','trump_gain')])
```

## Pennsylvania

```{r}
county_2012 %>% filter(state_name == 'Pennsylvania') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
pa_2012 <- (data_2012 %>% filter(state == 'Pennsylvania'))
kable(pa_2012[order(-pa_2012$trump_gain),c('county','trump_gain')])
```

## Wisconsin

```{r}
county_2012 %>% filter(state_name == 'Wisconsin') %>%
  ggplot(aes(long, lat, group = group, fill = trump_gain)) +
  geom_polygon(color = "#ffffff") +
  scale_fill_gradient2(labels = scales::percent,
                       low = muted('blue'),  mid = '#888888', na.value = 'gray80',
                       high = muted('red'), midpoint = 0,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Swing for Trump")
```

```{r}
wi_2012 <- (data_2012 %>% filter(state == 'Wisconsin'))
kable(wi_2012[order(-wi_2012$trump_gain),c('county','trump_gain')])
```

# Critical Result: Do these 2012 Obama-Trump counties matter? (Yes, they swung the election)

Do these counties really matter? If they did, they "swung" the election for Trump. Contrast this with a "surge" of voters in safe counties. Lets see if there are sufficient votes from these Obama-Trump counties to swing the election. First we need to get Clinton's results for the Obama-Trump counties:

```{r}
clinton_2016 <- relevant_data[relevant_data$state %in% counties_trump_won$state & relevant_data$county %in% counties_trump_won$county & relevant_data$party == 'democrat' & relevant_data$year == 2016,]
```

Now we merge the two datasets together to compute the margins in the Obama-Trump counties.

```{r}
merged_2016 <- merge(trump_intersect_2012, clinton_2016, by=c('state','county','FIPS'));
merged_2016$margin <- merged_2016$candidatevotes.x - merged_2016$candidatevotes.y

margin_by_counties_2016 <- merged_2016 %>% group_by(state) %>% summarize(margin = sum(margin))

margin_by_counties_2016$pivot <- 1 + (margin_by_counties_2016$margin %/% 2)

kable(margin_by_counties_2016)
```

We need to compare the results for the 7 states.

```{r}
state_2016 <- state_results[which(state_results$year == 2016 & state_results$state %in% merged_2016$state & (state_results$party == 'democrat' | state_results$party == 'republican')),]
```

We should also track the third party votes, since presumably they also acted as a [spoiler](https://en.wikipedia.org/wiki/Spoiler_effect).

```{r}
margin_2016_by_state <- state_2016 %>% group_by(state) %>% summarize(margin = max(candidatevotes) - min(candidatevotes), third_party_votes = max(totalvotes) - sum(candidatevotes))
kable(margin_2016_by_state)
```


Now we can answer the question, finally!

```{r}
pivots_by_state <- merge(margin_2016_by_state,margin_by_counties_2016,c('state'))
pivots_by_state$swingable <- pivots_by_state$margin.x < pivots_by_state$pivot
colnames(pivots_by_state)[colnames(pivots_by_state)=="margin.x"] <- "state_margin"
colnames(pivots_by_state)[colnames(pivots_by_state)=="margin.y"] <- "swing_counties_margin"
colnames(pivots_by_state)[colnames(pivots_by_state)=="third_party_votes.x"] <- "state_third_party_votes"
colnames(pivots_by_state)[colnames(pivots_by_state)=="third_party_votes.y"] <- "swing_counties_third_party_votes"
kable(pivots_by_state)
```
So, Iowa, Nebraska, and Ohio were not swingable. But if Florida, Michigan, Pennsylvania, and Wisconsin had swung for Clinton, that is a difference of 29, 16, 20, and 10 electoral votes for a sum of 75 electoral votes. Had this happen, Trump would have received 229 electoral votes to Clinton's 302.

On the other hand, the third party votes do not seem to credibly act as a spoiler.

## Florida Suspicions

There were two counties with Hacking suspicions in Florida. At first, Leon seems like a candidate, except it's the most highly educated county in the country (and there is a strong negative correlation between education and voting for Trump).

Madison county seems incredibly suspicious, it's approximately 40% African American and 60% white. Just from the Z-transform of vote percentage, this seems quite an amazing shift in votes for Clinton decreasing more than Trump's increase. (Break this out into its own post. Right now, just scratchwork.)

```{r}
fl_counties_pre_2016 <- county_results %>% filter(year < 2016 & state=='Florida')
fl_mean_sd <- fl_counties_pre_2016 %>% group_by(county,party) %>% summarize(mean=mean(votepercentage),sd=sd(votepercentage)) %>% filter(party=='democrat' | party=='republican')
fl_counties_2016 <- county_results %>% filter(year == 2016 & state=='Florida')
merged_fl <- merge(fl_counties_2016, fl_mean_sd, by=c('county','party'))
merged_fl$z <- (merged_fl$votepercentage - merged_fl$mean)/merged_fl$sd

kable(merged_fl[order(merged_fl$z),c('county','z','party')])
```

Well, what if this was a surge election, and it brought a lot of people out of the woodwork to vote for Trump? We can take this into account, normalizing statewide the county-'z' value to measure how variable each county was relative to the state as a whole:

```{r}
merged_fl$normed_z <- (merged_fl$z - mean(merged_fl$z))/sd(merged_fl$z)
kable(merged_fl[order(merged_fl$normed_z),c('county','normed_z','party')])
```

Remember: Leon county is the most educated in the country, it is not anomalous in defecting from Trump. The outliers have $|z|\geq 2$, so specifically Madison, DeSoto, and Glades counties appear statistically anomalous.

