---
title: "Most Surprising Counties"
author: "Alex Nelson"
date: "6/28/2019"
output: 
  md_document:
    variant: gfm
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(LaplacesDemon)
library(tidyverse)
library(rmarkdown)
knitr::opts_chunk$set(echo = TRUE)
```
```{r, include=FALSE}
source("../R/states.R")
source("../R/presidential_elections.R")
```


We can measure surprise using the Kullback–Leibler divergence, niftily provided by the `LaplacesDemon::KLD()` function. This quantifies how surprising the 2016 election results were, given we were assuming a result more closely resembling 2012.

```{r}
county_results <- make_party_into_factor(load_obj(county_path))
election_2004 <- filter(county_results, year == 2004)
election_2008 <- filter(county_results, year == 2008)
election_2012 <- filter(county_results, year == 2012)
election_2016 <- filter(county_results, year == 2016)
```

Now, we treat each county like a three-sided coin, with the proportion of the vote as a proxy for probability.

```{r}
conj_prob <- function(election) {
  prop <- election %>%
  group_by(year,state,county,party,FIPS) %>%
  transmute(probability = candidatevotes/totalvotes) %>%
  ungroup()
  prop$probability[is.na(prop$probability)] <- 0
  prop
}

prop_2004 <- conj_prob(election_2004)
prop_2008 <- conj_prob(election_2008)
prop_2012 <- conj_prob(election_2012)
prop_2016 <- conj_prob(election_2016)

probabilities2008 <- inner_join(select(prop_2004, -year),
                                select(prop_2008, -year),
                                by = c("state", "county", "party"))

probabilities2012 <- inner_join(select(prop_2008, -year),
                                select(prop_2012, -year),
                                by = c("state", "county", "party"))

probabilities2016 <- inner_join(select(prop_2012, -year),
                                select(prop_2016, -year),
                                by = c("state", "county", "party"))
```

We will now try to assign to each county the measure of surprise.

```{r}
compute_surprise <- function(data) {
  KLD(data$probability.x, data$probability.y)$sum.KLD.py.px
}

election_surprise <- function(probs) {
  probs %>%
    filter(party != "NA") %>%
    group_by(state,county) %>%
    nest() %>%
    mutate(surprise = map(data, compute_surprise)) %>%
    unnest(surprise, .drop = TRUE) %>%
    arrange(-surprise)
}

results08 <- election_surprise(probabilities2008)
results12 <- election_surprise(probabilities2012)
results16 <- election_surprise(probabilities2016)
```

## Statistically Significant Surprises

Note that Blair County, PA has a surprise of 0.01; did it vote differently than 2012 with any statistical significance (with alpha = 0.05, say):

```{r}
blair_pa_2012 <- election_2012 %>%
  filter(state == "Pennsylvania", county == "Blair", party == "republican")  %>%
  transmute(p = candidatevotes/totalvotes)

blair_pa_2016 <- election_2016 %>%
  filter(state == "Pennsylvania", county == "Blair", party == "republican")  %>%
  transmute(x = candidatevotes,
            n = totalvotes)

prop.test(blair_pa_2016$x, blair_pa_2016$n, blair_pa_2012$p)
```
The `p-value < 2.2e-16`, which tells us, yes, Blair county behaved differently in 2016 with statistical significance.

We could try generalizing this. Going county-by-county will be no good. We will instead work with state-level results, then examine each state for statistically different voting patterns among specifically Republicans.

```{r}
state_results <- make_party_into_factor(load_obj(state_path))
election_2012 <- filter(state_results, year == 2012)
election_2016 <- filter(state_results, year == 2016)
prop_2012 <- election_2012 %>%
  group_by(year,state,party) %>%
  transmute(probability = candidatevotes/totalvotes) %>%
  ungroup()
# prop_2012$probability[is.na(prop_2012$probability)] <- 0
```

We are interested in the situation where Trump did better than Republicans historically perform, as compared to the past race in 2012. This can be approximated using the p-value from the Binomial test.

```{r}
data <- inner_join(filter(election_2016, party == "republican", writein == F),
                   filter(prop_2012, party == "republican"),
                   by = c("state","party")) %>%
  filter(!is.na(totalvotes), is.state(state)) %>%
  select(state,candidatevotes,totalvotes,probability) %>%
  group_by(state) %>%
  mutate(trump_2016 = candidatevotes/totalvotes,
         p.value = binom.test(x = candidatevotes, n = totalvotes, p = probability, alternative = "greater")$p.value) %>%
  ungroup() %>%
  rename(republican_2012 = probability) %>%
  arrange(p.value)

kable(data)
```

There are onl 28 states where this p-value is smaller than units (of which 24 are really the interesting states). So we've narrowed down our attention to 24 states where Trump over-performed (compared to 2012). Really, there are only about a half dozen states worth investigating since they were won by Trump:

```{r}
kable(filter(data, trump_2016 > republican_2012, trump_2016 > 0.45))
```

## Picking Out the Relevant States

We can pick out the relevant states, but the criteria we have is exceedingly generous. We end up with 25 states, with genuine surprises like Michigan, but with false-positives like Alabama.

```{r}
relevant_states <- data %>%
  filter(p.value < 0.1) %>% 
  select(state) %>%
  `[[`("state") %>%
  sort
```

We could then compute the surprise for each of these states:

```{r}
relevant_surprise <- results16 %>%
  filter(state %in% relevant_states) %>%
  group_by(state) %>%
  summarize(total_surprise = sum(surprise))
```

Now we can filter out the states less surprising than Alabama:

```{r}
relevant_surprise <- filter(relevant_surprise,
                            total_surprise > relevant_surprise[which(relevant_surprise$state=="Alabama"),]$total_surprise) %>%
  arrange(-total_surprise)
```

This eliminates 10 false-positives. Lets try also weighing these states by electoral delegates:

```{r}
relevant_surprise %>%
  group_by(state) %>%
  mutate(ec = electoral_delegates[[state]],
         weighted_surprise = ec*total_surprise) %>%
  arrange(-weighted_surprise)
```
We shouldn't be surprised Nebraska was won by Trump, it's a red state... as are all the others with a small `weighted_surprise` value. So we can drop these as false-positives as well:

```{r}
true_surprises <- relevant_surprise %>%
  group_by(state) %>%
  mutate(electoral_delegates = electoral_delegates[[state]],
         weighted_surprise = electoral_delegates*total_surprise) %>%
  arrange(-weighted_surprise) %>%
  filter(weighted_surprise > 10)
```

The true surprises, where Trump overperformed, appears to be:

```{r}
kable(true_surprises)
```


## Overperformed...compared to what?

We should take the time here to note we should be a little careful upon further analysis to take, e.g., the running mean of the percentages for Republican candidates rather than just the last election's results.


